### 2011-05-13 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
updated and exported normalise method and vignette updates
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@55561 bc3139a8-67e5-0310-9ffc-ced21a209358

%\VignetteIndexEntry{Base Functions and Classes for MS-based Proteomics}
%\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Infrastructure, Bioinformatics, quantitative }
%\VignettePackage{MSnbase-demo}
\documentclass{article}
\usepackage{tikz}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

\SweaveOpts{echo=T,eval=T,fig=F,results=verbatim,cache=F,tikz=T,external=F,prefix.string=sweave-cache/figs/fig}

\newcommand{\R}{\texttt{R} }
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textit{#1}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

\hypersetup{%
  hyperindex = {true},
  colorlinks = {true},
  linktocpage = {true},
  plainpages = {false},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  pdfstartview = {Fit},
  pdfpagemode = {UseOutlines},
  pdfview = {XYZ null null null}
}

\input{Author.tex}

\begin{document}

\title{\Rpackage{MSnbase}: labelled MS2 data pre-processing, visualisation and quantification.}

\maketitle

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes the functionality implemented in the \Rpackage{MSnbase} package. 
  \Rpackage{MSnbase} aims at (1) facilitating the upload, processing, visualisation and 
  quantification of mass spectrometry data into the \R environment \citep{Rstat} by providing 
  specific data classes and methods and (2) enabling the utilisation of throughput-high data 
  analysis pipelines provided by the Bioconductor \citep{Gentleman2004} project. 
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, bioinformatics, quantitative.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

%% <<setup,cache=FALSE>>=
%%  setCacheDir("sweave-cache/values")
%% @

<<environment,cache=FALSE,echo=FALSE>>=
library("MSnbase")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Foreword.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}

\Rpackage{MSnbase} aims are providing a reproducible research framework to proteomics data analysis. 
It should allow researcher to easily mine mass spectrometry data, explore the data and its statistical properties 
and visually display these.

\Rpackage{MSnbase} also aims at being compatible with the infrastructure implemented in Bioconductor, 
in particular \Rpackage{Biobase}. As such, classes 
developed specifically for proteomics mass spectrometry data are based on the \Robject{eSet} 
and \Robject{Expression} classes. The main goal is to assure seamless compatibility with 
existing meta data structure, accessor methods and normalisation techniques.

This vignette illustrates \Rpackage{MSnbase} utility using a dummy data sets provided with the package 
without describing the underlying data structures. More details can be found in the 
package, classes, method and function documentations. A description of the classes is provided 
in the \texttt{MSnbase-development} vignette.

\input{NoteAboutSpeedAndMemory.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Data structure and content}\label{sec:data}

\subsection{Importing experiments}\label{sec:msnexp}

Raw data is contained in \Robject{MSnExp} objects, that stores all the spectra of an experiment, 
as defined by one or multiple raw data files.

\Rpackage{MSnbase} imports raw MS data stored in \texttt{mzXML} 
format \citep{Pedrioli2004} with the function \Rfunction{readMzXMLData} function. 
See \Rfunction{?readMzXMLData} for more details.
More formats, notably \texttt{mzML} \citep{Martens2010}, will be added at a later stage. 
Either MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
In this document, we will use a slightly bigger data set, \Robject{itraqdata}, provided with 
\Rpackage{MSnbase}. It includes feature metadata, accessible with the \Rfunction{fData} method. 
The metadata includes identification data for the \Sexpr{length(itraqdata)} MS2 spectra.

<<MSnExp,cache=FALSE,echo=TRUE>>=
library("MSnbase")
itraqdata
head(fData(itraqdata))
@

<<experiment-size,echo=FALSE,cache=FALSE>>=
sz <- sum(sapply(assayData(itraqdata),object.size)) + 
  object.size(itraqdata)
sz <- round(sz/(1024^2),2)
@

As illustrated above, showing the experiment textually displays it's content: 
\begin{itemize}
  \item Information about the raw data, i.e. the spectra.
  \item Specific information about the experiment processing\footnote{%%
    this part will be automatically updated when the object is 
    modified with it's \textit{ad hoc} methods, as illustrated later} and package version.
    This slot can be accessed with the \Rfunction{processingData} method.
  \item Other meta data, including experimental phenotype, 
    file name(s) used to import the data, protocol data, 
    information about features (individual spectra here) and
    experiment data. Most of these are implemented as 
    in the \Robject{eSet} class and are described in more details in their 
    respective manual pages. See \texttt{?MSnExp}
    and references therein for additional background information.

    The experiment meta data associated with an \Robject{MSnExp} experiment
    is of class \Robject{MIAPE}. It stores general information about the 
    experiment as well as MIAPE (Minimum Information About a Proteomics Experiment) 
    information \citep{Taylor2007, Taylor2008}.
    This meta-data can be accessed with the \Rfunction{experimentData} method. 
    When available, a summary of MIAPE-MS data can be printed with the \Rfunction{msInfo} method.
    See \texttt{?MIAPE} for more details.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spectra objects}\label{sec:spectra}

The raw data is composed of the \Sexpr{length(itraqdata)} MS spectra. 
The spectra are named individually 
(\Sexpr{paste(paste(head(featureNames(itraqdata)),collapse=", "),", ...",sep="")})
and stored in a \Robject{environment}. They can be accessed individually with 
\texttt{itraqdata[["X1"]]} or \texttt{itraqdata[[1]]}, 
or as a list with \texttt{spectra(itraqdata)}.
As we have loaded our experiment specifying \texttt{msLevel=2}, 
the spectra will all be of level 2 (or higher, if available).

<<Spectrum,cache=FALSE,echo=TRUE>>=
sp <- itraqdata[["X1"]]
sp
@

Attributes of individual spectra or of all spectra of an experiment can be 
accessed with their respective methods: \Rfunction{precursorCharge} for the 
precursor charge, \Rfunction{rtime} for the retention time, \Rfunction{mz} 
for the MZ values, \Rfunction{intensity} for the intensities, \ldots see the 
\Robject{Spectrum}, \Robject{Spectrum1} and \Robject{Spectrum2} manuals for more details.

<<accessors,cache=FALSE,echo=TRUE>>=
peaksCount(sp)
head(peaksCount(itraqdata))
rtime(sp)
head(rtime(itraqdata))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Reporter ions}\label{sec:reporterions}

Reporter ions are defined with the \Robject{ReporterIons} class. 
Specific peaks of interest are defined by a MZ value, a with 
around the expected MZ and a name (and optionally a colour for plotting, 
see section \ref{sec:plotting}). \Robject{ReporterIons} instances 
are required to quantify reporter peaks in \Robject{MSnExp} 
experiments. Instances for the most commonly used isobaric tags like 
iTRAQ 4-plex and 8-plex and TMT tags are already defined in 
\Rpackage{MSnbase}. See \texttt{?ReporterIons} for details about how 
to generate new \Robject{ReporterIons} objects.

<<ReporterIons>>=
iTRAQ4
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Plotting raw data}\label{sec:plotting}

Spectra can be plotted individually or as part of (subset) experiments 
with the \Robject{plot} method. Full spectra can be plotted 
(using \texttt{full=TRUE}), specific reporter ions of interest 
(by specifying with reporters with \texttt{reporters=iTRAQ4} for instance) 
or both (see figure \ref{fig:spectrum-plot}).

\begin{figure}[!ht]
<<spectrum-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=5>>=
plot(sp,reporters=iTRAQ4,full=TRUE) 
@ 
\caption{Raw MS2 spectrum with details about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

<<bsasel,echo=FALSE>>=
bsasel <- fData(itraqdata)$ProteinAccession=="BSA"
@

It is also possible to plot all spectra of an experiment (figure \ref{fig:msnexp-plot}). 
Lets start by subsetting the \Robject{itraqdata} experiment using the protein accession 
numbers included in the feature metadata, and keep the \Sexpr{sum(bsasel)} from 
the \textit{BSA} protein.

<<subset,echo=TRUE>>=
sel <- fData(itraqdata)$ProteinAccession=="BSA"
bsa <- itraqdata[sel]
bsa
as.character(fData(bsa)$ProteinAccession)
@

%% Lets start by extracting all spectra that have the same precursor MZ value than \Robject{sp} 
%% into a separate experiment, using the \Rfunction{extractPrecSpectra} method. 
%% This method takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.
%% <<extractPrecSpec,echo=T,cache=F>>=
%% exp2 <- extractPrecSpectra(raw.experiment,precursorMz(sp))
%% exp2
%% @

These can then be visualised together by plotting the 
\Robject{MSnExp} object, as illustrated on figure \ref{fig:msnexp-plot}.

\begin{figure}[!ht]
<<msnexp-plot,fig=T,echo=T,tikz=T,external=F,width=5,height=6>>=
plot(bsa,reporters=iTRAQ4,full=FALSE)
@ 
\caption{Experiment-wide raw MS2 spectra. The y-axes of the individual 
  spectra are automatically rescaled to the same range. 
  See section \ref{sec:norm} to rescale peaks identically. }
\label{fig:msnexp-plot}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data processing}\label{sec:processing}

There are several methods implemented to perform basic data manipulation. 
Low intensity peaks can be set to 0 with the \Rfunction{removePeaks} method 
from spectra or whole experiments. The intensity threshold below which peaks are removed 
is defined by the \Robject{t} parameter. \Robject{t} can be specified directly 
as a numeric. The default value is the character \texttt{"min"}, that will remove 
all peaks equal to the lowest non null intensity in any spectrum. 
We observe the effect of the \Rfunction{removePeaks} method by comparing 
total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{tic} method
before (object \Robject{itraqdata}) and after (object \Robject{experiment}) for 
spectrum \texttt{X55}.

<<removePeaks,echo=TRUE,cache=FALSE>>=
experiment <- removePeaks(itraqdata,t=400,verbose=FALSE)
## total ion current
tic(itraqdata[["X55"]])
tic(experiment[["X55"]])
@

<<prepare-spectrum-plot2,echo=FALSE>>=
p1 <- plot(itraqdata[["X55"]],full=TRUE) + theme_gray(5) 
p2 <- plot(experiment[["X55"]],full=TRUE) + theme_gray(5) 
@

\begin{figure}[!ht]
<<spectrum-plot2,fig=T,echo=F,tikz=T,external=F,width=5,height=2.5>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p1,vp=vplayout(1,1))
print(p2,vp=vplayout(1,2))
@ 
\caption{Same spectrum before (left) and after setting peaks <= 400 to 0.}
\label{fig:spectrum-plot2}
\end{figure}

Unlike the name might suggest, the \Rfunction{removePeaks} method does 
not actually remove peaks from the spectrum; they are set to 0. This 
can be checked using the \Rfunction{peaksCount} method, that returns 
the number of peaks (including 0 intensity peaks) in a spectrum.
To effectively remove 0 intensity 
peaks from spectra, and reduce the size of the data set, one can use 
the \Rfunction{clean} method. 
The effect of the \Rfunction{removePeaks} and \Rfunction{clean} methods 
are illustrated on figure \ref{fig:preproc} on page \pageref{fig:preproc}.

<<clean,echo=TRUE,cache=FALSE>>=
## number of peaks
peaksCount(itraqdata[["X55"]])
peaksCount(experiment[["X55"]])
experiment <- clean(experiment,verbose=FALSE)
peaksCount(experiment[["X55"]])
@

<<preprosp,cache=FALSE,echo=FALSE>>=
int <- c(0,1,1,3,1,1,0,0,0,1,3,7,3,1,0)
mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25,
        114.3,114.35,114.4,114.42,114.48,114.5,114.55,114.6)
ppsp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
p1 <- plot(ppsp,full=TRUE) + theme_gray(5) + 
  geom_point(size=3,alpha=I(1/3)) +
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Original spectrum") 
p2 <- plot(removePeaks(ppsp,t=3),full=TRUE) + 
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) + 
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Peaks < 3 removed") 
p3 <- plot(clean(removePeaks(ppsp,t=3)),full=TRUE) + 
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) + 
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Peaks < 3 removed and cleaned") 
@

\begin{figure}[p]
  \begin{center}
<<preproc-plot,fig=T,echo=F,tikz=T,external=F,width=3,height=6>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(3, 1)))
print(p1, vp=vplayout(1,1))
print(p2, vp=vplayout(2,1))
print(p3, vp=vplayout(3,1))
@ 
\caption{This figure illustrated the effect or the \Rfunction{removePeaks} 
  and \Rfunction{clean} methods. The left-most spectrum displays two peaks, 
  of max height 3 and 7 respectively. The middle spectrum shows the result 
  of calling \Rfunction{removePeaks} with argument \Robject{t=3}, which sets 
  all data points of the first peak, whose maximum height is smaller or equal 
  to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean} 
  after \Rfunction{removePeaks} effectively deletes successive 0 intensities from 
  the spectrum, as shown on the right plot. }
\label{fig:preproc}
  \end{center}
\end{figure}

Another useful manipulation method is \Rfunction{trimMz}, that takes as 
parameters and \Robject{MSnExp} (or a \Robject{Spectrum}) and a numeric 
\Robject{mzlim}. MZ values smaller then \texttt{min(mzlim)} 
or greater then \texttt{max(mzmax)} are discarded. This method is 
particularly useful when one wants to concentrate on a specific MZ range, 
as for reporter ions quantification, and generally results in substantial 
reduction of data size. Compare the size of the full trimmed experiment to 
the original \Sexpr{sz} Mb.
\label{trimMz-example}

<<trimMz,echo=TRUE,cache=FALSE>>=
range(mz(itraqdata[["X55"]]))
experiment <- trimMz(experiment,mzlim=c(112,120))
range(mz(experiment[["X55"]]))
experiment
@ 
As can be seen above, all processing performed on the experiment is recorded 
and displayed as integral part of the experiment object.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{From spectra to quantitative expression data}

\subsection{Reporter ions quantitation}\label{sec:quant}

Quantitation is performed on fixed peaks in the spectra, that are specified 
with an \Robject{ReporterIons} object. A specific peak is defined by it's expected 
\Robject{mz} value and is searched for within \Robject{mz} $\pm$ \Robject{width}.  
If no data is found, \Robject{NA} is returned.

<<reporters,echo=TRUE,cache=FALSE>>=
mz(iTRAQ4)
width(iTRAQ4)
@

The \Rfunction{quantify} method takes the following parameters: 
an \Robject{MSnExp} experiment, 
a character describing the quantification \Robject{method}, 
the \Robject{reporters} to be quantified 
and a \Robject{strict} logical defining whether data points 
ranging outside of \Robject{mz} $\pm$ \Robject{width} should be considered 
for quantitation. 
Additionally, a progress bar can be displaying when setting 
the \Robject{verbose} parameter to \Robject{TRUE}.
Three quantification methods are implemented, as illustrated 
on figure \ref{fig:quant-methods}: 
\Robject{trapezoidation} returns the area under the peak of interest, 
\Robject{max} returns the apex of the peak and 
\Robject{sum} returns the sum of all intensities of the peak. 
See \texttt{?quantify} for more details.

<<simplesp,cache=FALSE,echo=FALSE>>=
int <- c(0,1,1,3,1,1,0)
mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25)
ssp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
p <- plot(ssp,full=TRUE)
p <- p + theme_gray(5)
@

\begin{figure}[p]
<<quantitation-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=5>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p + opts(title="Quantitation using 'sum'") + geom_point(size=3,alpha=I(1/3),colour="red"), vp=vplayout(1,1))
print(p + opts(title="Quantitation using 'max'") + geom_point(aes(x=114.1,y=3),alpha=I(1/3),colour="red",size=3), vp=vplayout(1,2))
print(p + opts(title="Trapezoidation and strict=FALSE") + geom_polygon(alpha=I(1/5),fill="red"),vp=vplayout(2,1))
print(p + opts(title="Trapezoidation and strict=TRUE") + 
      geom_polygon(aes(x=c(114.05,114.05,114.1,114.15,114.15),y=c(0,1,3,1,0)),fill="red",alpha=I(1/5)),vp=vplayout(2,2))
@ 
\caption{The different quantitation methods are illustrated above. 
  Quantitation using \Robject{sum} sums all the data points in the peaks to 
  produce, for this example, \Sexpr{quantify(ssp,iTRAQ4[1],method="sum")[[1]]}, 
  whereas method \Robject{max} only uses the peak's maximum intensity, 
  \Sexpr{quantify(ssp,iTRAQ4[1],method="max")[[1]]}. 
  \Robject{Trapezoidation} calculates the area under the peak taking 
  the full with into account (using \Robject{strict=FALSE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=FALSE)[[1]],3)}) 
  or only the width as defined by the reporter 
  (using \Robject{strict=TRUE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=TRUE)[[1]],3)}). }
\label{fig:quant-methods}
\end{figure}

\bigskip

The \Rfunction{quantify} method returns \Robject{MSnSet} objects, that 
extend the well-known \Robject{eSet} class defined in the 
\Rpackage{Biobase} package. \Robject{MSnSet} instances are very similar 
to \Robject{ExpressionSet} objects, except for the experiment meta-data 
that captures MIAPE specific information.
The assay data is a matrix of dimensions 
$n \times m$, where $m$ is the  number of features/spectra 
originally in the \Robject{MSnExp} used as parameter in 
\Robject{quantify} and $m$ is the number of reporter ions, that 
can be accessed with the \Rfunction{exprs} method.
The meta data is directly inherited from the \Robject{MSnExp} instance.

<<quantify,echo=TRUE,cache=FALSE,keep.source=T>>=
qnt <- quantify(experiment,
                method="trap",
                reporters=iTRAQ4,
                strict=FALSE,
                verbose=FALSE)
qnt
head(exprs(qnt))
@

If no peak is detected for a reporter ion peak, the respective 
quantitation value is set to \texttt{NA}. In our case, there 
is \Sexpr{sum(is.na(exprs(qnt)))} such case in row 
\Sexpr{which(is.na(exprs(qnt))) %% nrow(qnt)}.
We will remove the offending line, as we do not expect any missing  peaks.

<<removeNa,echo=TRUE>>=
sum(is.na(exprs(qnt)))
whichRow <- which(is.na(exprs(qnt))) %% nrow(qnt)
qnt <- qnt[-whichRow,]
sum(is.na(exprs(qnt)))
@

\subsection{Normalisation}\label{sec:norm}

A \Robject{MSnSet} object is meant to be compatible with further downstream 
packages for data normalisation and statistical analysis. There is also a 
\Rfunction{normalise} method for expression sets. The method takes and instance 
of class \Robject{MSnSet} as first argument, and a character to describe the 
\Robject{method} to be used:

\begin{description}
\item{\Robject{quantiles}} Applies quantile normalisation \citep{Bolstad03} as 
  implemented in the \Rfunction{normalize.quantiles} function of the 
  \Rpackage{preprocessCore} package. 
\item{\Robject{quantiles.robust}} Applies robust quantile normalisation 
  \citep{Bolstad03} as implemented in the 
  \Rfunction{normalize.quantiles.robust} function of the 
  \Rpackage{preprocessCore} package. 
\item{\Robject{vsn}} Applies variance stabilisation normalization \citep{Huber2002} 
  as implemented in the \Rfunction{vsn2} function of the 
  \Rpackage{vsn} package. 
\item{\Robject{max}} Each feature's reporter intensity is divided by the 
  maximum of the reporter ions intensities.
\item{\Robject{sum}} Each feature's reporter intensity is divided by the 
  sum of the reporter ions intensities.
\end{description}

<<normalise,echo=TRUE,cache=FALSE>>=
qnt.max <- normalise(qnt,"max")
qnt.sum <- normalise(qnt,"sum")
qnt.quant <- normalise(qnt,"quantiles")
qnt.qrob <- normalise(qnt,"quantiles.robust")
qnt.vsn <- normalise(qnt,"vsn")
@

The effect of these are illustrated on figure \ref{fig:norm-plot} and
figure \ref{fig:cv-plot} reproduces figure 3 of \cite{Karp2010} that described 
the application of vsn on iTRAQ reporter data.

\begin{figure}[p]
<<norm-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=7>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[6],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt,ttl="Non-normalised data")
.plot(qnt.max)
.plot(qnt.sum)
.plot(qnt.quant)
.plot(qnt.qrob)
.plot(qnt.vsn)
@ 
\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}

<<cvdata,echo=FALSE,cache=FALSE>>=
sd1 <- apply(log2(exprs(qnt))+10,1,sd)
mn1 <- apply(log2(exprs(qnt))+10,1,mean)
cv1 <- sd1/mn1
sd2 <- apply(exprs(qnt.vsn)+10,1,sd)
mn2 <- apply(exprs(qnt.vsn)+10,1,mean)
cv2 <- sd2/mn2
dfr <- rbind(data.frame(rank=order(mn1),cv=cv1,norm="raw"),
             data.frame(rank=order(mn2),cv=cv2,norm="vsn"))
library("zoo")
## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
## rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
##
## Calling directly rollapply.zoo to make it zoo_1.6-4 compatible.
## The above requires zoo >= 1.7-0, which is as of 15 March 2011
## not yet available on CRAN (only on r-forge).
rmed1 <- zoo:::rollapply.zoo(cv1,7,function(x) median(x,na.rm=TRUE))
rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
dfr2 <- 
  rbind(data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed1,norm="raw"),
        data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed2,norm="vsn"))

p <- qplot(rank,cv,data=dfr,col=norm) + 
  geom_line(data=dfr2,aes(x=x,y=y,colour=norm)) +
  theme_gray(7)
@

\begin{center}
\begin{figure}[!ht]
<<cv-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=4>>=
print(p)
@ 
\caption{CV versus signal intensity comparison for log2 and vsn
  transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}
\end{center}

Note that it is also possible to normalise individual spectra or whole experiments with 
the \Rfunction{normalise} method using the \Robject{max} method. 
This will rescale all peaks between 0 and 1. 
To visualise the relative reporter peaks, one should this first trim the spectra 
using method \Rfunction{trimMz} as illustrated in section \ref{sec:processing}, 
then normalise the \Robject{MSnExp} with \Rfunction{normalise} using 
\Robject{method="max"} as illustrated above and plot the data using \Rfunction{plot} 
(figure \ref{fig:msnexp-norm-plot}).

<<prepare-msnset-nrom-plot,cache=FALSE,echo=FALSE>>=
p <- plot(normalise(experiment[bsasel],"max"),reporters=iTRAQ4,full=FALSE)
p <- p + theme_gray(7)
@

\begin{figure}[p]
<<msnexp-norm-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=6>>=
print(p)
@ 
\caption{Experiment-wide normalised MS2 spectra. The y-axes of the individual 
  spectra is now rescaled between 0 and 1 (highest peak), as opposed to 
  figure \ref{fig:msnexp-plot}. }
\label{fig:msnexp-norm-plot}
\end{figure}

\subsection{Feature aggregation}
\label{sec:feataggregation}

The above quantitation and normalisation has been performed on quantitative data 
obtained from individual spectra. However, the biological unit of interest is not 
the spectrum but the peptide or the protein. As such, it is important to be able to 
summarise features that belong to a same group, i.e. spectra from one peptide, 
peptides that originate from one protein, or directly combine all spectra that have 
been uniquely associated to one protein.

\Rpackage{MSnbase} provides one function, \Rfunction{combineFeatures}, that allows 
to aggregate features stored in an \Robject{MSnSet} using build-in or user 
defined summary function and return a new \Robject{MSnSet} instance. 
The three main arguments are described below. 
Additional details can be found in the method documentation.

<<makeGroups1,echo=FALSE,cache=FALSE>>=
gb <- fData(qnt)$ProteinAccession
@

\Rfunction{combineFeatures}'s first argument, \Robject{object}, is an instance of class \Robject{MSnSet}, 
as has been created in the section \ref{sec:quant} for instance. 
The second argument, \Robject{groupBy}, is a \Robject{factor} than has as many elements as 
there are features in the \Robject{MSnSet} \Robject{object} argument. The features corresponding 
to the \Robject{groupBy} levels will be aggregated so that the resulting \Robject{MSnSet} output 
will have \Rfunction{length(levels(groupBy))} features. Here, we will combine individual 
MS2 spectra based on the protein they originate from. 
As shown below, this will result in \Sexpr{length(table(gb))} new aggregated features.

<<makeGroups2,echo=TRUE,cache=FALSE>>=
gb <- fData(qnt)$ProteinAccession
table(gb)
length(unique(gb))
@

The third argument, \Robject{fun}, 
defined how to combine the features. Predefined functions are readily available and can be 
specified as strings 
(\Robject{fun="mean"}, \Robject{fun="median"}, \Robject{fun="sum"}, 
\Robject{fun="weighted.mean"} or \Robject{fun="medianpolish"} to compute respectively the mean, 
media, sum, weighted mean or median polish of the features to be aggregated). 
Alternatively, is is possible to supply user defined functions with 
\Robject{fun=function(x) \{ ... \}}. We will use the \Robject{median} here.

<<combineFeatures,echo=TRUE,cache=FALSE>>=
qnt2 <- combineFeatures(qnt,groupBy=gb,fun="median")
qnt2
@

%% Other intended ways to define aggregation groups is to import 
%% identification data into R and collate it to the \Robject{MSnSet} 
%% feature data, and subsequently use peptide sequences or protein assignments 
%% to combine features.

\subsection{Purity correction}
\label{sec:purcor}

Impurities in the reporter reagents can bias the results and can be corrected when 
manufacturers provide correction coefficients. 
These generally come as percentages of each reporter ion that have masses differing 
by -2, -1, +1 and +2 Da from the nominal reporter ion mass due to isotopic variants.  
The \Rfunction{purityCorrect} method applies such correction to \Robject{MSnSet} 
instances. It also requires a square matrix as second argument, \Robject{impurities}, 
that defines the relative percentage of reporter in the quantified each peak. 
See \texttt{?purityCorrect} for more details.

<<purityCorrect,echo=TRUE,cache=FALSE>>=
impurities <- matrix(c(0.929,0.059,0.002,0.000,
                       0.020,0.923,0.056,0.001,
                       0.000,0.030,0.924,0.045,
                       0.000,0.001,0.040,0.923),
                     nrow=4)
qnt2.crct <- purityCorrect(qnt2,impurities)
head(exprs(qnt2))
head(exprs(qnt2.crct))
@

\subsection{Label-free MS2 quantitation}
\label{sec:spectralcounting}

Note that if samples are not multiplexed, label-free MS2 quantitation is possible 
using \Rpackage{MSnbase}.
Once individual spectra have been assigned to peptides and proteins, it becomes 
straightforward to estimate protein quantities using the spectral counting method, 
as illustrated in section \ref{sec:feataggregation}, when the \Robject{groupBy} 
argument is defined.

\section{Quantitative assessment of incomplete dissociation}
\label{sec:incompdissoc}

Quantitation using isobaric reporter tags assumes complete dissociation between
the reporter group (red on figure \ref{fig:itraqchem}), 
balance group (blue) 
and peptide (the peptide reactive group is drawn in green). 
However, incomplete dissociation does occur and 
results in an isobaric tag (i.e reporter and balance groups) specific peaks.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=4cm]{./itraqchem.pdf}
\caption{ iTRAQ 4-plex isobaric tags reagent consist of three parts: 
  (1) a charged reporter group (MZ of 114, 115, 116 and 117) that is unique to each of the four reagents (red),
  (2) an uncharged mass balance group (28-31 Da) (blue)and 
  (3) a peptide reactive group (NHS ester) that binds to the peptide. 
  In case of incomplete dissociation, the reporter and balance groups produce a specific peaks 
  at MZ 145.
}
\label{fig:itraqchem}
  \end{center}
\end{figure}

\Rpackage{MSnbase} provides, among others, a \Robject{ReporterIons} object for iTRAQ 4-plex that includes the 
145 peaks, called \Robject{iTRAQ5}. This can then be used to quantify the experiment as show in section \ref{sec:quant} 
to estimate incomplete dissociation for each spectrum.

<<incompdiss,echo=TRUE,cache=FALSE,keep.source=T>>=
iTRAQ5
incompdiss <- quantify(itraqdata,
                       method="trap",
                       reporters=iTRAQ5,
                       strict=FALSE,
                       verbose=FALSE)
head(exprs(incompdiss))
@

Figure \ref{fig:incompdiss} compares these intensities for the whole experiment.

\begin{figure}[!hbt]
  \begin{center}
<<incompdiss-plot,fig=T,echo=F,tikz=T,external=F,width=4,height=3.5>>=
dfr <- melt(exprs(incompdiss))
colnames(dfr) <- c("feature","reporters","intensity")
qplot(x=reporters,y=intensity,data=dfr,geom="boxplot") + theme_gray(8)
@ 
\caption{ Boxplot comparing intensities of the 4 reporter ions an the 
  incomplete dissociation specific peak. }
\label{fig:incompdiss}
  \end{center}
\end{figure}

%% \section{Conclusion}\ref{sec:ccl}

\clearpage

\section{Session information}\label{sec:sessionInfo} 
<<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
toLatex(sessionInfo())
@

\bibliographystyle{plainnat}
\bibliography{MSnbase}

\end{document}