### 2011-09-20 inst/doc/Foreword.tex
Spectrum2 ms1scan slot updated to precScanNum, using mzR::header()$precursorScanNum to populate. Updated and exported readIspyData.
 o removed (internal) Mascot query link column in readIspyData 
   to work with latest ouput version <2011-09-12 Mon>
 o removed the fillUp part in readIspyData <2011-09-14 Wed>
 o exported readIspyData <2011-09-20 Tue>
 o removed link to proteomics sig list <2011-09-20 Tue>
 o removed url in DESCRIPTION <2011-09-20 Tue>
 o Spectrum2 slot ms1scan renamed to precScanNum and populated using 
   mzR::header()$precursorScanNum. Updated affected methods/functions 
   a

\section*{Foreword}

\Rpackage{MSnbase} is in an early development %%(package version \Sexpr{packageVersion("MSnbase")}, 
(see section \ref{sec:sessionInfo} for details about packages and version used in this vignette). 
Although main data structures have been thought out and are meant to be compatible with other existing 
mature infrastructure in the Bioconductor project, changes may occur in the future. 
Current functionality will evolve and new one will be added. 
Although at an early stage, this package is released in the hope that it may foster  
new developments in proteomics data analysis within \R by providing a common infrastructure. 
Several package developers working with mass spectrometry and proteomics data met at the 
Bioconductor Developer Meeting Europe\footnote{\url{http://bioconductor.org/help/course-materials/2010/HeidelbergNovember2010/}} 
held in Heidelberg in November 2010, and agreed to combine efforts. 
This library is one attempt to do so.

\bigskip

You are welcome to contact me for questions, bugs, typos or suggestions about \Rpackage{MSnbase}. 
If you wish to reach a broader audience for general questions about proteomics analysis using \texttt{R}, 
you may want to use the Bioconductor mailing 
list\footnote{\url{https://stat.ethz.ch/mailman/listinfo/bioconductor}}.
%% proteomics special interest 
%% group\footnote{\url{https://stat.ethz.ch/mailman/listinfo/bioc-sig-proteomics}}.