### 2011-09-28 inst/doc/MSnbase-io/MSnbase-io.Rnw
new vignette files
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@58520 bc3139a8-67e5-0310-9ffc-ced21a209358

%\VignetteIndexEntry{MSnbase IO capabilities}
%\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
%\VignettePackage{MSnbase-io}
\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,shadows,fit}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

\SweaveOpts{echo=T,eval=T,fig=F,results=verbatim,cache=F,tikz=T,external=T,prefix.string=sweave-cache/figs/fig}

\newcommand{\R}{\texttt{R} }
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

\hypersetup{%
  hyperindex = {true},
  colorlinks = {true},
  linktocpage = {true},
  plainpages = {false},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  pdfstartview = {Fit},
  pdfpagemode = {UseOutlines},
  pdfview = {XYZ null null null}
}

\input{Author.tex}

\begin{document}

%% pgf settup
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
% Define block styles
\tikzstyle{input} = [rectangle, draw, fill=blue!20,
  text width=6em, text centered, rounded corners, minimum height=4em]
\tikzstyle{fun} = [rectangle, draw, fill=white, drop shadow, 
  text width=7em, text centered, rounded corners, minimum height=2em]
\tikzstyle{obj} = [rectangle, draw, fill=red!20, 
  text width=5em, text centered, rounded corners, minimum height=5em]

\title{\Rpackage{MSnbase} input/output capabilities}

\maketitle

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes \Rpackage{MSnbase}'s input and output 
  capabilites.
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, IO.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

%% <<setup,cache=FALSE>>=
%% setCacheDir("sweave-cache/values")
%% @

%% <<environment,cache=FALSE,echo=FALSE>>=
%% library(MSnbase)
%% @

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\input{Foreword.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}

\Rpackage{MSnbase}'s aims are to facilitate the reproducible analysis of 
mass spectrometry data within the \R environment, from raw data import and processing, 
feature quantification, quantification and statistical analysis of the results.
Several input functions are provided to allow execution of all of several 
of these steps. Intermediate or final results can also be saved or exported. 
These capabilites are presented below.

\section{Data input}

\paragraph{Raw data} Data stored in one of the published \texttt{XML}-based 
formats. i.e. \texttt{mzXML} \citep{Pedrioli2004}, \texttt{mzData} \citep{Orchard2007} 
or \texttt{mzML} \citep{Martens2010}, 
can be imported with the \Rfunction{readMSData} method, which makes use of 
the \Rpackage{mzR} package to create \Robject{MSnExp} objects. 
The files can be in profile or centroided mode. 
See \Rfunction{?readMSData} for details.

\paragraph{Peak lists} Peak lists in the \texttt{mgf} 
format\footnote{\url{http://www.matrixscience.com/help/data\_file\_help.html\#GEN}} 
can be imported using the \Rfunction{readMgfData}. 
In this case, the peak data has generally been pre-processed by other software.
See \Rfunction{?readMgfData} for details.

\paragraph{Quantitation data} Third party software can be used to generate 
quantitative data and exported as a spreadsheet (generally comma or tab separated format). 
This data as well as any additional metadata can be imported with the 
\Rfunction{readMSnSet} function. See \Rfunction{?readMSnSet} for details.

\begin{figure}[!htb] %% input
  \begin{center}
    \begin{tikzpicture}[node distance = 2cm, auto]
      % Place nodes
      \node [input] (raw) {Raw data in and open \texttt{XML} format};
      \node [fun, left of=raw, node distance=4cm] (readMSData) {\Rfunction{readMSData}};
      \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
      \node [fun, left of=mgf, node distance=4cm] (readMgfData) {\Rfunction{readMgfData}};
      \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
      \node [fun, left of=spreadsheet, node distance=4cm] (readMSnSet) {\Rfunction{readMSnSet}};
      \node [obj, left of=readMSData, yshift=-2.5mm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
      \node [obj, left of=readMSnSet, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
      % Background
      \begin{pgfonlayer}{background}
        \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
      \end{pgfonlayer}
      % Draw edges
      \draw  (raw) -- (readMSData);
      \draw  (mgf) -- (readMgfData);
      \draw  (spreadsheet) -- (readMSnSet);
      \draw [->] (readMSData) -- (MSnExp);
      \draw [->] (readMgfData) -- (MSnExp);
      \draw [->] (readMSnSet) -- (MSnSet);
      \draw  (MSnExp) -- (quantify);
      \draw [->] (quantify) -- (MSnSet);
    \end{tikzpicture}
    \caption{Illustration of \texttt{MSnbase} input capabilities. 
      The white and red boxes represent \R functions/methods and objects respectively. 
      The blue boxes represent different disk storage formats.}
    \label{fig:input}
  \end{center}
\end{figure}

\section{Data output}

\paragraph{RData files} \R objects can most easily be stored on disk with the \Rfunction{save} function.
It creates compressed binary images of the data representation that can later be read back from the file 
with the \Rfunction{load} function.

\paragraph{Peak lists} \Robject{MSnExp} instances as well as individual spectra can be written 
as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the metadata in the original 
\R object can not be included in the file. See \Rfunction{?writeMgfData} for details.

\paragraph{Quantitation data} Quantitation data can be exported to spreadsheet files with 
the \Rfunction{write.exprs} method. Feature metadata can be appended to the feature 
intenstiy values. See \Rfunction{?writeMgfData} for details.

\begin{figure}[!htb] %% out
  \begin{center}
    \begin{tikzpicture}[node distance = 2cm, auto]
      % Place nodes
      \node [input] (raw) {Raw data in and open \texttt{XML} format};
      \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
      \node [fun, right of=mgf, node distance=4cm, yshift=1cm] (writeMgfData) {\Rfunction{writeMgfData}};
      \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
      \node [fun, right of=spreadsheet, node distance=4cm] (writeexprs) {\Rfunction{write.exprs}};

      \node [obj, right of=writeMgfData, yshift=1.25cm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
      \node [obj, right of=writeexprs, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
      % Background
      \begin{pgfonlayer}{background}
        \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
      \end{pgfonlayer}
      % Draw edges
      \draw  (MSnExp) -- (writeMgfData);
      \draw [->] (writeMgfData) -- (mgf);
      \draw (MSnSet) -- (writeexprs);
      \draw [->] (writeexprs) -- (spreadsheet);
      \draw  (MSnExp) -- (quantify);
      \draw [->] (quantify) -- (MSnSet);
    \end{tikzpicture}
    \caption{Illustration of \texttt{MSnbase} output capabilities.
      The white and red boxes represent \R functions/methods and objects respectively. 
      The blue boxes represent different disk storage formats.}
    \label{fig:output}
  \end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \section{Session information}
%% \label{sec:sessionInfo}

%% <<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
%% toLatex(sessionInfo())
%% @

\bibliographystyle{plainnat}
\bibliography{MSnbase}

\end{document}