### 2012-08-14 inst/doc/MSnbase-io/MSnbase-io.Rnw

%\VignetteIndexEntry{MSnbase IO capabilities}
%\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
%\VignettePackage{MSnbase-io}
\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,shadows,fit}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

\newcommand{\R}{\texttt{R} }
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

\usepackage{geometry}
\geometry{verbose,
  tmargin = 2.5cm,
  bmargin = 2.5cm,
  lmargin = 3.0cm,
  rmargin = 3.0cm}

\usepackage{hyperref}
\usepackage{breakurl}
\hypersetup{%
  pdfusetitle,
  bookmarks = {true},
  bookmarksnumbered = {true},
  bookmarksopen = {true},
  bookmarksopenlevel = 2,
  unicode = {true},
  breaklinks = {false},
  hyperindex = {true},
  colorlinks = {true},
  linktocpage = {true},
  plainpages = {false},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  pdfstartview = {Fit},
  pdfpagemode = {UseOutlines},
  pdfview = {XYZ null null null}
}

%% pgf setup
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
% Define block styles
\tikzstyle{input} = [rectangle, draw, fill=blue!20,
  text width=6em, text centered, rounded corners, minimum height=4em]
\tikzstyle{fun} = [rectangle, draw, fill=white, drop shadow, 
  text width=7em, text centered, rounded corners, minimum height=2em]
\tikzstyle{obj} = [rectangle, draw, fill=red!20, 
  text width=5em, text centered, rounded corners, minimum height=5em]

\input{Author.tex}

\begin{document}

\title{\Rpackage{MSnbase} input/output capabilities}

\maketitle

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes \Rpackage{MSnbase}'s input and output 
  capabilites.
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, IO.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% \tableofcontents

\vspace{10mm}

<<knitr, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(fig.align = 'center', 
               fig.show = 'hold', 
               par = TRUE,
               prompt = TRUE,
               comment = NA)
options(replace.assign = TRUE, 
        width = 70)
knit_hooks$set(par = function(before, options, envir) {
  if (before && options$fig.show != 'none') 
    par(mar = c(4,4,.1,.1),
        cex.lab = .95,
        cex.axis = .9,
        mgp = c(2,.7,0),
        tcl = -.3)
})
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\input{Foreword.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}

\Rpackage{MSnbase}'s aims are to facilitate the reproducible analysis of 
mass spectrometry data within the \R environment, from raw data import and processing, 
feature quantification, quantification and statistical analysis of the results.
Data import functions for several formats are provided and intermediate or final 
results can also be saved or exported. 
These capabilites are presented below.

\section{Data input}

\paragraph{Raw data} Data stored in one of the published \texttt{XML}-based 
formats. i.e. \texttt{mzXML} \citep{Pedrioli2004}, \texttt{mzData} \citep{Orchard2007} 
or \texttt{mzML} \citep{Martens2010}, 
can be imported with the \Rfunction{readMSData} method, which makes use of 
the \Rpackage{mzR} package to create \Robject{MSnExp} objects. 
The files can be in profile or centroided mode. 
See \Rfunction{?readMSData} for details.

\paragraph{Peak lists} Peak lists in the \texttt{mgf} 
format\footnote{\url{http://www.matrixscience.com/help/data\_file\_help.html\#GEN}} 
can be imported using the \Rfunction{readMgfData}. 
In this case, the peak data has generally been pre-processed by other software.
See \Rfunction{?readMgfData} for details.

\paragraph{Quantitation data} Third party software can be used to generate 
quantitative data and exported as a spreadsheet (generally comma or tab separated format). 
This data as well as any additional metadata can be imported with the 
\Rfunction{readMSnSet} function. See \Rfunction{?readMSnSet} for details.

\Rpackage{MSnbase} also supports the \texttt{mzTab} 
format\footnote{\url{http://code.google.com/p/mztab/}}, a light-weight, tab-delimited 
file format for proteomics data developed within the Proteomics Standards Initiative (PSI).
\texttt{mzTab} files can be read into \R with \Rfunction{readMzTabData} to create and \Robject{MSnSet} instance.

\begin{figure}[!htb] %% input
  \begin{center}
    \begin{tikzpicture}[node distance = 2cm, auto]
      % Place nodes
      \node [input] (raw) {Raw data in an open \texttt{XML} format};
      \node [fun, left of=raw, node distance=4cm] (readMSData) {\Rfunction{readMSData}};
      \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
      \node [fun, left of=mgf, node distance=4cm] (readMgfData) {\Rfunction{readMgfData}};
      \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
      \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
      \node [fun, left of=spreadsheet, node distance=4cm] (readMSnSet) {\Rfunction{readMSnSet}};
      \node [fun, left of=mztab, node distance=4cm] (readMzTabData) {\Rfunction{readMzTabData}};
      \node [obj, left of=readMSData, yshift=-2.5mm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
      \node [obj, left of=readMSnSet, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
      % Background
      \begin{pgfonlayer}{background}
        \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
      \end{pgfonlayer}
      % Draw edges
      \draw  (raw) -- (readMSData);
      \draw  (mgf) -- (readMgfData);
      \draw  (spreadsheet) -- (readMSnSet);
      \draw [->] (readMSData) -- (MSnExp);
      \draw [->] (readMgfData) -- (MSnExp);
      \draw [->] (readMSnSet) -- (MSnSet);
      \draw  (MSnExp) -- (quantify);
      \draw [->] (quantify) -- (MSnSet);
      \draw (mztab) -- (readMzTabData);
      \draw [->] (readMzTabData) -- (MSnSet);
    \end{tikzpicture}
    \caption{Illustration of \texttt{MSnbase} input capabilities. 
      The white and red boxes represent \R functions/methods and objects respectively. 
      The blue boxes represent different disk storage formats.}
    \label{fig:input}
  \end{center}
\end{figure}

\section{Data output}

\paragraph{RData files} \R objects can most easily be stored on disk with the \Rfunction{save} function.
It creates compressed binary images of the data representation that can later be read back from the file 
with the \Rfunction{load} function.

\paragraph{Peak lists} \Robject{MSnExp} instances as well as individual spectra can be written 
as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the metadata in the original 
\R object can not be included in the file. See \Rfunction{?writeMgfData} for details.

\paragraph{Quantitation data} Quantitation data can be exported to spreadsheet files with 
the \Rfunction{write.exprs} method. Feature metadata can be appended to the feature 
intenstiy values. See \Rfunction{?writeMgfData} for details.

\Robject{MSnSet} instances can also be exported to \texttt{mzTab} files using the 
\Rfunction{writeMzTabData} function.

\begin{figure}[!htb] %% out
  \begin{center}
    \begin{tikzpicture}[node distance = 2cm, auto]
      % Place nodes
      \node [input] (raw) {Raw data in an open \texttt{XML} format};
      \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
      \node [fun, right of=mgf, node distance=4cm, yshift=1cm] (writeMgfData) {\Rfunction{writeMgfData}};
      \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
      \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
      \node [fun, right of=spreadsheet, node distance=4cm] (writeexprs) {\Rfunction{write.exprs}};
      \node [fun, right of=mztab, node distance=4cm] (writeMzTabData) {\Rfunction{writeMzTabData}};
      \node [obj, right of=writeMgfData, yshift=1.25cm, node distance=4cm] (MSnExp) {\Robject{MSnExp}};
      \node [obj, right of=writeexprs, node distance=4cm] (MSnSet) {\Robject{MSnSet}};
      \node [fun, below of=MSnExp] (quantify) {\Rfunction{quantify}};
      % Background
      \begin{pgfonlayer}{background}
        \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
      \end{pgfonlayer}
      % Draw edges
      \draw  (MSnExp) -- (writeMgfData);
      \draw [->] (writeMgfData) -- (mgf);
      \draw (MSnSet) -- (writeexprs);
      \draw [->] (writeexprs) -- (spreadsheet);
      \draw  (MSnExp) -- (quantify);
      \draw [->] (quantify) -- (MSnSet);
      \draw (MSnSet) -- (writeMzTabData);
      \draw [->] (writeMzTabData) -- (mztab);
    \end{tikzpicture}
    \caption{Illustration of \texttt{MSnbase} output capabilities.
      The white and red boxes represent \R functions/methods and objects respectively. 
      The blue boxes represent different disk storage formats.}
    \label{fig:output}
  \end{center}
\end{figure}

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \section{Session information}
%% \label{sec:sessionInfo}

%% <<label=sessioninfo,results=tex,echo=FALSE,cache=FALSE>>=
%% toLatex(sessionInfo())
%% @

\bibliographystyle{plainnat}
\bibliography{MSnbase}

\end{document}