### 2012-08-14 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
using knitr instead of pgfSweave and misc vignette updated
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@68442 bc3139a8-67e5-0310-9ffc-ced21a209358

%\VignetteIndexEntry{Base Functions and Classes for MS-based Proteomics}
%\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Infrastructure, Bioinformatics, quantitative }
%\VignettePackage{MSnbase-demo}
\documentclass{article}
\usepackage{tikz}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

\newcommand{\R}{\texttt{R} }
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

\usepackage{geometry}
\geometry{verbose,
  tmargin = 2.5cm,
  bmargin = 2.5cm,
  lmargin = 3.5cm,
  rmargin = 3.5cm}

\usepackage{hyperref}
\usepackage{breakurl}
\hypersetup{%
  pdfusetitle,
  bookmarks = {true},
  bookmarksnumbered = {true},
  bookmarksopen = {true},
  bookmarksopenlevel = 2,
  unicode = {true},
  breaklinks = {false},
  hyperindex = {true},
  colorlinks = {true},
  linktocpage = {true},
  plainpages = {false},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  pdfstartview = {Fit},
  pdfpagemode = {UseOutlines},
  pdfview = {XYZ null null null}
}

\input{Author.tex}

\begin{document}

\title{\Rpackage{MSnbase}: labelled MS2 data pre-processing, visualisation and quantification.}

\maketitle

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes the functionality implemented in the \Rpackage{MSnbase} package. 
  \Rpackage{MSnbase} aims at (1) facilitating the import, processing, visualisation and 
  quantification of mass spectrometry data into the \R environment \citep{Rstat} by providing 
  specific data classes and methods and (2) enabling the utilisation of throughput-high data 
  analysis pipelines provided by the Bioconductor \citep{Gentleman2004} project. 
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, bioinformatics, quantitative.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

<<environment, cache=FALSE, echo=FALSE>>=
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("zoo"))
suppressPackageStartupMessages(library("doMC"))
suppressPackageStartupMessages(library("multicore"))
library("grid")
library("reshape2")
@

<<knitr, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(fig.align = 'center', 
               fig.show = 'hold', 
               par = TRUE,
               prompt = TRUE,
               comment = NA)
options(replace.assign = TRUE, 
        width = 68)
knit_hooks$set(par = function(before, options, envir) {
  if (before && options$fig.show != 'none') 
    par(mar = c(4,4,.1,.1),
        cex.lab = .95,
        cex.axis = .9,
        mgp = c(2,.7,0),
        tcl = -.3)
})
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Foreword.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}

\Rpackage{MSnbase} \citep{Gatto2012} aims are providing a reproducible research framework to proteomics data analysis. 
It should allow researcher to easily mine mass spectrometry data, explore the data and its statistical properties 
and visually display these.

\Rpackage{MSnbase} also aims at being compatible with the infrastructure implemented in Bioconductor, 
in particular \Rpackage{Biobase}. As such, classes 
developed specifically for proteomics mass spectrometry data are based on the \Robject{eSet} 
and \Robject{Expression} classes. The main goal is to assure seamless compatibility with 
existing meta data structure, accessor methods and normalisation techniques.

This vignette illustrates \Rpackage{MSnbase} utility using a dummy data sets provided with the package 
without describing the underlying data structures. More details can be found in the 
package, classes, method and function documentations. A description of the classes is provided 
in the \texttt{MSnbase-development} vignette.

\input{NoteAboutSpeedAndMemory.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Data structure and content}\label{sec:data}

\subsection{Importing experiments}\label{sec:io}

\Rpackage{MSnbase} is able to import raw MS data stored in one of the \texttt{XML}-based formats 
as well as peak lists in the \texttt{mfg} format\footnote{%%
Mascot Generic Format -- \url{http://www.matrixscience.com/help/data\_file\_help.html\#GEN}}

\paragraph{Raw data} The \texttt{XML}-based formats, \texttt{mzXML} \citep{Pedrioli2004}, 
\texttt{mzData} \citep{Orchard2007} and \texttt{mzML} \citep{Martens2010} can be imported 
with the \Rfunction{readMSData} function, as illstrated below 
(see \Rfunction{?readMSData} for more details).

<<readdata, echo=TRUE, cache=FALSE>>=
file <- dir(system.file(package = "MSnbase", dir = "extdata"),
            full.names = TRUE, pattern = "mzXML$")
rawdata <- readMSData(file, msLevel = 2, verbose = FALSE)
@

Either MS1 or MS2 spectra can be loaded at a time by setting the \texttt{msLevel} parameter accordingly.
In this document, we will use the \Robject{itraqdata} data set, provided with 
\Rpackage{MSnbase}. It includes feature metadata, accessible with the \Rfunction{fData} accessor. 
The metadata includes identification data for the \Sexpr{length(itraqdata)} MS2 spectra.

\paragraph{Peak lists} Peak lists can often be exported after spectrum processing 
from vendor-specific software and are also used as input to search engines. 
Peak lists in \texttt{mgf} format can be imported with the function 
\Rfunction{readMgfData} (see \Rfunction{?readMgfData} for details)
to create experiment objects. 
Experiments or individual spectra can be exported to an \texttt{mgf} file with 
the \Rfunction{writeMgfData} methods (see \Rfunction{?writeMgfData} for details and examples).

\bigskip

\paragraph{Experiments with multiple runs} Although it is possible to load and process multiple 
files serially and later merge the resulting quantitation data as show in section 
\ref{sec:combine} (page \pageref{sec:combine}), it is also feasible to load several raw data files at once. 
Here, we report the analysis of an LC-MSMS experiment were 
14 liquid chromatography (LC) fractions were loaded using \Rfunction{readMSData} on a 
32-cores servers with 128 Gb of RAM. 
It took about 90 minutes to read the 14 uncentroided \texttt{mzXML} raw files (4.9 Gb on disk in total)
and create a 3.3 Gb raw data object (an \Robject{MSnExp} instance, see next section). 
Quantitation of 9 reporter ions (\Robject{iTRAQ9} object, see \ref{sec:reporterions}) 
for 88690 features was performed in parallel on 16 processors and took 76 minutes. 
The resulting quantitation data was only 22.1 Mb and could easily be further processed and 
analysed on a standard laptop computer.

\bigskip

See also section \ref{sec:io2} to import quantitative data stored in spreadsheets into \R 
for further processing using \Rpackage{MSnbase}. The \texttt{MSnbase-io} vignette 
gives a general overview of \Rpackage{MSnbase}'s input/ouput capabilites.

\subsection{MS experiments}\label{sec:msnexp}

Raw data is contained in \Robject{MSnExp} objects, that stores all the spectra of an experiment, 
as defined by one or multiple raw data files.

<<MSnExp,cache=FALSE,echo=TRUE>>=
library("MSnbase")
itraqdata
head(fData(itraqdata))
@

<<experiment-size,echo=FALSE,cache=FALSE>>=
sz <- sum(sapply(assayData(itraqdata),object.size)) + 
  object.size(itraqdata)
sz <- round(sz/(1024^2),2)
@

As illustrated above, showing the experiment textually displays it's content: 
\begin{itemize}
  \item Information about the raw data, i.e. the spectra.
  \item Specific information about the experiment processing\footnote{%%
    this part will be automatically updated when the object is 
    modified with it's \textit{ad hoc} methods, as illustrated later} and package version.
    This slot can be accessed with the \Rfunction{processingData} method.
  \item Other meta data, including experimental phenotype, 
    file name(s) used to import the data, protocol data, 
    information about features (individual spectra here) and
    experiment data. Most of these are implemented as 
    in the \Robject{eSet} class and are described in more details in their 
    respective manual pages. See \texttt{?MSnExp}
    and references therein for additional background information.

    The experiment meta data associated with an \Robject{MSnExp} experiment
    is of class \Robject{MIAPE}. It stores general information about the 
    experiment as well as MIAPE (Minimum Information About a Proteomics Experiment) 
    information \citep{Taylor2007, Taylor2008}.
    This meta-data can be accessed with the \Rfunction{experimentData} method. 
    When available, a summary of MIAPE-MS data can be printed with the \Rfunction{msInfo} method.
    See \texttt{?MIAPE} for more details.
\end{itemize}

\subsection{Spectra objects}\label{sec:spectra}

The raw data is composed of the \Sexpr{length(itraqdata)} MS spectra. 
The spectra are named individually 
(\Sexpr{paste(paste(head(featureNames(itraqdata)),collapse=", "),", ...",sep="")})
and stored in a \Robject{environment}. They can be accessed individually with 
\texttt{itraqdata[["X1"]]} or \texttt{itraqdata[[1]]}, 
or as a list with \texttt{spectra(itraqdata)}.
As we have loaded our experiment specifying \texttt{msLevel=2}, 
the spectra will all be of level 2 (or higher, if available).

<<Spectrum,cache=FALSE,echo=TRUE>>=
sp <- itraqdata[["X1"]]
sp
@

Attributes of individual spectra or of all spectra of an experiment can be 
accessed with their respective methods: \Rfunction{precursorCharge} for the 
precursor charge, \Rfunction{rtime} for the retention time, \Rfunction{mz} 
for the MZ values, \Rfunction{intensity} for the intensities, ... see the 
\Robject{Spectrum}, \Robject{Spectrum1} and \Robject{Spectrum2} manuals for more details.

<<accessors,cache=FALSE,echo=TRUE>>=
peaksCount(sp)
head(peaksCount(itraqdata))
rtime(sp)
head(rtime(itraqdata))
@

\subsection{Reporter ions}\label{sec:reporterions}

Reporter ions are defined with the \Robject{ReporterIons} class. 
Specific peaks of interest are defined by a MZ value, a with 
around the expected MZ and a name (and optionally a colour for plotting, 
see section \ref{sec:plotting}). \Robject{ReporterIons} instances 
are required to quantify reporter peaks in \Robject{MSnExp} 
experiments. Instances for the most commonly used isobaric tags like 
iTRAQ 4-plex and 8-plex and TMT tags are already defined in 
\Rpackage{MSnbase}. See \texttt{?ReporterIons} for details about how 
to generate new \Robject{ReporterIons} objects.

<<ReporterIons>>=
iTRAQ4
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Plotting raw data}\label{sec:plotting}

\subsection{Default plots}\label{sec:defaultplots}

Spectra can be plotted individually or as part of (subset) experiments 
with the \Robject{plot} method. Full spectra can be plotted 
(using \texttt{full=TRUE}), specific reporter ions of interest 
(by specifying with reporters with \texttt{reporters=iTRAQ4} for instance) 
or both (see figure \ref{fig:spectrum-plot}).

\begin{figure}[!ht]
<<spectrum-plot, dev='tikz', fig.width=4, fig.height=4, fig.keep='high'>>=
plot(sp, reporters = iTRAQ4, full = TRUE) 
@ 
\caption{Raw MS2 spectrum with details about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

<<bsasel, echo=FALSE>>=
bsasel <- fData(itraqdata)$ProteinAccession == "BSA"
@

It is also possible to plot all spectra of an experiment (figure \ref{fig:msnexp-plot}). 
Lets start by subsetting the \Robject{itraqdata} experiment using the protein accession 
numbers included in the feature metadata, and keep the \Sexpr{sum(bsasel)} from 
the \textit{BSA} protein.

<<subset,echo=TRUE>>=
sel <- fData(itraqdata)$ProteinAccession == "BSA"
bsa <- itraqdata[sel]
bsa
as.character(fData(bsa)$ProteinAccession)
@

%% Lets start by extracting all spectra that have the same precursor MZ value than \Robject{sp} 
%% into a separate experiment, using the \Rfunction{extractPrecSpectra} method. 
%% This method takes an \Robject{MSnExp} experiment and a precursor MZ value as parameters.
%% <<extractPrecSpec,echo=T,cache=F>>=
%% exp2 <- extractPrecSpectra(raw.experiment,precursorMz(sp))
%% exp2
%% @

These can then be visualised together by plotting the 
\Robject{MSnExp} object, as illustrated on figure \ref{fig:msnexp-plot}.

\begin{figure}[p]
<<msnexp-plot, dev='tikz', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
@ 
\caption{Experiment-wide raw MS2 spectra. The y-axes of the individual 
  spectra are automatically rescaled to the same range. 
  See section \ref{sec:norm} to rescale peaks identically. }
\label{fig:msnexp-plot}
\end{figure}

\subsection{Customising your plots}\label{sec:customplots}

The \Rpackage{MSnbase} \Rfunction{plot} methods have a logical \Robject{plot} parameter 
(default is \Robject{TRUE}), that specifies if the plot should be printed to the current 
device. A plot object is also (invisibly) returned, so that it can be saved as a variable 
for later use or for customisation.

\Rpackage{MSnbase} uses the \Rpackage{ggplot2} package to generate figures, 
which can subsequently easily be customised. 
More details about \Rpackage{ggplot2} can be found in \cite{ggplot2} (especially chapter 8) 
and on \href{http://had.co.nz/ggplot2/}{http://had.co.nz/ggplot2/}. 
Finally, if a plot object has been saved in a variable \Robject{p}, 
it is possible to obtain a summary about the object with \Rfunction{summary(p)}. 
To view the data frame used to generate the plot, use \Rfunction{p@data}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quality control}\label{sec:qc}

The current section is not executed dynamically for package size and processing time constrains. 
The figures and tables have been generated with the respective methods and included statically 
in the vignette for illustration purposes.

\bigskip

\Rpackage{MSnbase} allows easy and flexible access to the data, which allows to 
visualise data features to assess it's quality. Some methods are readily available, 
although many QC approaches will be experiment specific and users are encourage to 
explore their data.

The \Rfunction{plot2d} method takes one \Robject{MSnExp} instance as firs argument 
to produce retention time \textit{vs.} precursor MZ scatter plots. Points represent 
individual MS2 spectra and can be coloured based on 
precursor charge (with second argument \Robject{z="charge"}), 
total ion count (\Robject{z="ionCount"}),
number of peaks in the MS2 spectra \Robject{z="peaks.count"}) or, 
when multiple data files were loaded, file \Robject{z="file"}), 
as illustrated on figure \ref{fig:plot2d}. 
The lower right panel is produced for only a subset of proteins. 
See the method documentation for more details.

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{./plot2d-figure.png}
\caption{ Illustration of the \Rfunction{plot2d} output. } 
\label{fig:plot2d}
\end{figure}

The \Rfunction{plotDensity} method illustrates the distribution 
of several parameters of interest (see figure \ref{fig:plotDensity}).
Similarly to \Rfunction{plot2d}, the first argument is an \Robject{MSnExp} instance. 
The second is one of \Robject{precursor.mz}, \Robject{peaks.count} or \Robject{ionCount}, 
whose density will be plotted. 
An optional third argument specifies whether the x axes should be logged.

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{./plotDensity-figure.png}
\caption{ Illustration of the \Rfunction{plotDensity} output. } 
\label{fig:plotDensity}
\end{figure}

%% xtable(preprocSelectionTable(velos))
%% % latex table generated in R 2.14.0 by xtable 1.5-6 package
%% % Thu May 19 10:22:29 2011
%% \begin{table}[ht]
%% \begin{center}
%% \begin{tabular}{rr}
%%   \hline
%%  & x \\ 
%%   \hline
%% 1 & 6836 \\ 
%%   2 &  59 \\ 
%%   3 &   5 \\ 
%%    \hline
%% \end{tabular}
%% \end{center}
%% \end{table}

%% matplot on iTRAQ5 quantitation (see allquant code chunk)

\bigskip

The \Rfunction{plotMzDelta} method\footnote{The code to generate the 
  histograms has been contributed by Guangchuang Yu from Jinan University, 
  China.} implements the M/Z delta plot from \cite{Foster11} 
The M/Z delta plot illustrates the suitability of MS2 spectra for
identification by plotting the M/Z differences of the most intense
peaks. The resulting histogram should optimally shown outstanding
bars at amino acid residu masses. More details and parameters 
are described in the method documentation (\Rfunction{?plotMzDelta}).
Figure \ref{fig:plotMzDelta-pride12011} has been generated using 
the PRIDE experiment 12011, as in \cite{Foster11}.

\begin{figure}[htb]
  \begin{center}   
    \includegraphics[width=.7\linewidth]{./plotMzDelta-pride12011.pdf}
\caption{ Illustration of the \Rfunction{plotMzDelta} output for the PRIDE experiment 12011, 
  as in figure 4A from \cite{Foster11}. } 
\label{fig:plotMzDelta-pride12011}
  \end{center}
\end{figure}

\bigskip

In section \ref{sec:incompdissoc} on page \pageref{sec:incompdissoc}, 
we illustrate how to assess incomplete reporter ion dissociation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data processing}\label{sec:processing}

\subsection{Cleaning spectra}\label{sec:clean}

There are several methods implemented to perform basic data manipulation. 
Low intensity peaks can be set to 0 with the \Rfunction{removePeaks} method 
from spectra or whole experiments. The intensity threshold below which peaks are removed 
is defined by the \Robject{t} parameter. \Robject{t} can be specified directly 
as a numeric. The default value is the character \texttt{"min"}, that will remove 
all peaks equal to the lowest non null intensity in any spectrum. 
We observe the effect of the \Rfunction{removePeaks} method by comparing 
total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{ionCount} 
method before (object \Robject{itraqdata}) and after (object \Robject{experiment}) for 
spectrum \texttt{X55}. The respective spectra are shown on 
figure \ref{fig:spectrum-clean-plot} (page \pageref{fig:spectrum-clean-plot}).

<<removePeaks,echo=TRUE,cache=FALSE>>=
experiment <- removePeaks(itraqdata, t = 400, verbose = FALSE)
## total ion current
ionCount(itraqdata[["X55"]])
ionCount(experiment[["X55"]])
@

\begin{figure}[!ht]
<<spectrum-clean-plot, dev='tikz', fig.width=6, fig.height=3, echo=FALSE, fig.keep='high'>>=
p1 <- plot(itraqdata[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
p2 <- plot(experiment[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p1,vp=vplayout(1,1))
print(p2,vp=vplayout(1,2))
@ 
\caption{Same spectrum before (left) and after setting peaks <= 400 to 0.}
\label{fig:spectrum-clean-plot}
\end{figure}

Unlike the name might suggest, the \Rfunction{removePeaks} method does 
not actually remove peaks from the spectrum; they are set to 0. This 
can be checked using the \Rfunction{peaksCount} method, that returns 
the number of peaks (including 0 intensity peaks) in a spectrum.
To effectively remove 0 intensity 
peaks from spectra, and reduce the size of the data set, one can use 
the \Rfunction{clean} method. 
The effect of the \Rfunction{removePeaks} and \Rfunction{clean} methods 
are illustrated on figure \ref{fig:preproc} on page \pageref{fig:preproc}.

<<clean, echo=TRUE, cache=FALSE>>=
## number of peaks
peaksCount(itraqdata[["X55"]])
peaksCount(experiment[["X55"]])
experiment <- clean(experiment, verbose = FALSE)
peaksCount(experiment[["X55"]])
@

<<preprosp, cache=FALSE, echo=FALSE>>=
int <- c(0,1,1,3,1,1,0,0,0,1,3,7,3,1,0)
mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25,
        114.3,114.35,114.4,114.42,114.48,114.5,114.55,114.6)
ppsp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
p1 <- plot(ppsp, full = TRUE, plot = FALSE) + theme_gray(5) + 
  geom_point(size=3,alpha=I(1/3)) +
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Original spectrum") 
p2 <- plot(removePeaks(ppsp,t=3), full=TRUE, plot = FALSE) + 
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) + 
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Peaks < 3 removed") 
p3 <- plot(clean(removePeaks(ppsp,t=3)), full = TRUE, plot = FALSE) + 
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) + 
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Peaks < 3 removed and cleaned") 
@

\begin{figure}[p]
<<preproc-plot, dev='tikz', fig.width=3, fig.height=6, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(3, 1)))
print(p1, vp=vplayout(1,1))
print(p2, vp=vplayout(2,1))
print(p3, vp=vplayout(3,1))
@ 
\caption{This figure illustrated the effect or the \Rfunction{removePeaks} 
  and \Rfunction{clean} methods. The left-most spectrum displays two peaks, 
  of max height 3 and 7 respectively. The middle spectrum shows the result 
  of calling \Rfunction{removePeaks} with argument \Robject{t=3}, which sets 
  all data points of the first peak, whose maximum height is smaller or equal 
  to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean} 
  after \Rfunction{removePeaks} effectively deletes successive 0 intensities from 
  the spectrum, as shown on the right plot. }
\label{fig:preproc}
\end{figure}

\subsection{Focusing on specific MZ values}\label{sec:trim}

Another useful manipulation method is \Rfunction{trimMz}, that takes as 
parameters and \Robject{MSnExp} (or a \Robject{Spectrum}) and a numeric 
\Robject{mzlim}. MZ values smaller then \texttt{min(mzlim)} 
or greater then \texttt{max(mzmax)} are discarded. This method is 
particularly useful when one wants to concentrate on a specific MZ range, 
as for reporter ions quantification, and generally results in substantial 
reduction of data size. Compare the size of the full trimmed experiment to 
the original \Sexpr{sz} Mb.
\label{trimMz-example}

<<trimMz,echo=TRUE,cache=FALSE>>=
range(mz(itraqdata[["X55"]]))
experiment <- trimMz(experiment, mzlim = c(112,120))
range(mz(experiment[["X55"]]))
experiment
@ 
As can be seen above, all processing performed on the experiment is recorded 
and displayed as integral part of the experiment object.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{From spectra to quantitative expression data}

\subsection{Reporter ions quantitation}\label{sec:quant}

Quantitation is performed on fixed peaks in the spectra, that are specified 
with an \Robject{ReporterIons} object. A specific peak is defined by it's expected 
\Robject{mz} value and is searched for within \Robject{mz} $\pm$ \Robject{width}.  
If no data is found, \Robject{NA} is returned.

<<reporters,echo=TRUE,cache=FALSE>>=
mz(iTRAQ4)
width(iTRAQ4)
@

The \Rfunction{quantify} method takes the following parameters: 
an \Robject{MSnExp} experiment, 
a character describing the quantification \Robject{method}, 
the \Robject{reporters} to be quantified 
and a \Robject{strict} logical defining whether data points 
ranging outside of \Robject{mz} $\pm$ \Robject{width} should be considered 
for quantitation. 
Additionally, a progress bar can be displaying when setting 
the \Robject{verbose} parameter to \Robject{TRUE}.
Three quantification methods are implemented, as illustrated 
on figure \ref{fig:quant-methods}: 
\Robject{trapezoidation} returns the area under the peak of interest, 
\Robject{max} returns the apex of the peak and 
\Robject{sum} returns the sum of all intensities of the peak. 
See \texttt{?quantify} for more details.

<<simplesp, cache=FALSE, echo=FALSE, fig.keep='none'>>=
int <- c(0,1,1,3,1,1,0)
mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25)
ssp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
p <- plot(ssp, full=TRUE, plot=FALSE)
p <- p + theme_gray(5)
@

\begin{figure}[p]
<<quantitation-plot, dev='tikz',fig.width=5, fig.height=5, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p + opts(title = "Quantitation using 'sum'") +
      geom_point(size = 3, alpha = I(1/3), colour = "red"),
      vp = vplayout(1, 1))
print(p + opts(title = "Quantitation using 'max'") +
      geom_point(aes(x = 114.1, y = 3),
                 alpha = I(1/18),
                 colour = "red", size = 3),
      vp = vplayout(1, 2))
print(p +
      opts(title = "Trapezoidation and strict=FALSE") +
      geom_polygon(alpha = I(1/5), fill = "red"),
      vp = vplayout(2, 1))
print(p +
      opts(title = "Trapezoidation and strict=TRUE") + 
      geom_polygon(aes(x = c(NA, 114.05, 114.05, 114.1, 114.15, 114.15, NA),
                       y=c(NA, 0, 1, 3, 1, 0, NA)), fill = "red", alpha = I(1/5)),
      vp = vplayout(2,2))

@ 
\caption{The different quantitation methods are illustrated above. 
  Quantitation using \Robject{sum} sums all the data points in the peaks to 
  produce, for this example, \Sexpr{quantify(ssp,iTRAQ4[1],method="sum")[[1]]}, 
  whereas method \Robject{max} only uses the peak's maximum intensity, 
  \Sexpr{quantify(ssp,iTRAQ4[1],method="max")[[1]]}. 
  \Robject{Trapezoidation} calculates the area under the peak taking 
  the full with into account (using \Robject{strict=FALSE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=FALSE)[[1]],3)}) 
  or only the width as defined by the reporter 
  (using \Robject{strict=TRUE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=TRUE)[[1]],3)}). }
\label{fig:quant-methods}
\end{figure}

\bigskip

The \Rfunction{quantify} method returns \Robject{MSnSet} objects, that 
extend the well-known \Robject{eSet} class defined in the 
\Rpackage{Biobase} package. \Robject{MSnSet} instances are very similar 
to \Robject{ExpressionSet} objects, except for the experiment meta-data 
that captures MIAPE specific information.
The assay data is a matrix of dimensions 
$n \times m$, where $m$ is the  number of features/spectra 
originally in the \Robject{MSnExp} used as parameter in 
\Robject{quantify} and $m$ is the number of reporter ions, that 
can be accessed with the \Rfunction{exprs} method.
The meta data is directly inherited from the \Robject{MSnExp} instance.

<<quantify,echo=TRUE,cache=FALSE,keep.source=T>>=
qnt <- quantify(experiment,
                method = "trap",
                reporters = iTRAQ4,
                strict = FALSE,
                parallel = FALSE,
                verbose = FALSE)
qnt
head(exprs(qnt))
@

If no peak is detected for a reporter ion peak, the respective 
quantitation value is set to \texttt{NA}. In our case, there 
is \Sexpr{sum(is.na(exprs(qnt)))} such case in row 
\Sexpr{which(is.na(exprs(qnt))) %% nrow(qnt)}.
We will remove the offending line, as we do not expect any missing peaks.

<<removeNa,echo=TRUE>>=
table(is.na(qnt))
whichRow <- which(is.na((qnt))) %% nrow(qnt)
qnt <- qnt[-whichRow, ]
sum(is.na(exprs(qnt)))
@

The \Rfunction{filterNA} method allows to filter features based on the proportion 
of missing data their carry, specified using the \Robject{pNA} argument. 
The same result as above could have been achieved with 
\Rfunction{qnt <- filterNA(qnt, pNA = 0)}, 
specifying that we allow 0 percent of missing data. 
See also the \Rfunction{plotNA} method to get an overview of the completeness 
of a data set.

\subsection{Importing quantitation data}\label{sec:io2}

If quantitation data is already available as a spreadsheet, it can be 
imported, along with additional optional feature and sample (pheno) meta data, 
with the \Rfunction{readMSnSet} function. This function takes the 
respective text-based spreadsheet (comma- or tab-separated) file names 
as argument to create a valid \Robject{MSnSet} instance.

Note that the quantitation data of \Robject{MSnSet} objects can also be 
exported to a text-based spreadsheet file using the \Rfunction{write.exps} 
method.

\Rpackage{MSnbase} also supports the \texttt{mzTab} 
format\footnote{\url{http://code.google.com/p/mztab/}}, a light-weight, tab-delimited 
file format for proteomics data. \texttt{mzTab} files can be read into \R with 
\Rfunction{readMzTabData} to create and \Robject{MSnSet} instance. 
\Robject{MSnSet} objects can also be exported to \texttt{mzTab} with the 
\Rfunction{writeMzTabData} function.

See the \texttt{MSnbase-io} vignette for a general overview of 
\Rpackage{MSnbase}'s input/ouput capabilites.

\subsection{Peak adjustments}\label{sec:purcor}

\paragraph{Single peak adjustment} In certain cases, peak intensities need to be adjusted 
as a result of peak interferance. For example, the $+1$ peak of the phenylalanine (F, Phe) 
immonium ion (with m/z 120.03) inteferes with the 121.1 TMT reporter ion. 
Below, we calculate the relative intensity of the +1 peaks compared to the main peak using the 
\Rpackage{Rdispo} package.

<<pheplus1, echo=TRUE, cache=FALSE>>=
library(Rdisop)
## Phenylalanine immonium ion
Fim <- getMolecule("C8H10N")
getMass(Fim)
isotopes <- getIsotope(Fim)
F1 <- isotopes[2, 2]
F1
@

If desired, one can thus specifically quantify the F immonium ion in the MS2 spectrum, 
estimate the intensity of the +1 ion (\Sexpr{round(F1,4)}\% of the F peak) and substract 
this calculated value from the 121.1 TMT reporter intensity.

The above principle can also be generalised for a set of overlapping peaks, 
as described below.

\paragraph{Reporter ions purity correction} Impurities in the reporter reagents can also 
bias the results and can be corrected when manufacturers provide correction coefficients. 
These generally come as percentages of each reporter ion that have masses differing 
by -2, -1, +1 and +2 Da from the nominal reporter ion mass due to isotopic variants.  
The \Rfunction{purityCorrect} method applies such correction to \Robject{MSnSet} 
instances. It also requires a square matrix as second argument, \Robject{impurities}, 
that defines the relative percentage of reporter in the quantified each peak. 
See \texttt{?purityCorrect} for more details.

<<purityCorrect,echo=TRUE,cache=FALSE>>=
impurities <- matrix(c(0.929, 0.059, 0.002, 0.000,
                       0.020, 0.923, 0.056, 0.001,
                       0.000, 0.030, 0.924, 0.045,
                       0.000, 0.001, 0.040, 0.923),
                     nrow = 4)
qnt.crct <- purityCorrect(qnt, impurities)
head(exprs(qnt))
head(exprs(qnt.crct))
@

The \Rfunction{makeImpuritiesMatrix} can be used to create impurity matrices. 
It opens a rudimentary spreadsheet that can be directly edited.

\subsection{Normalisation}\label{sec:norm}

A \Robject{MSnSet} object is meant to be compatible with further downstream 
packages for data normalisation and statistical analysis. There is also a 
\Rfunction{normalise} method for expression sets. The method takes and instance 
of class \Robject{MSnSet} as first argument, and a character to describe the 
\Robject{method} to be used:

\begin{description}
\item{\Robject{quantiles}} Applies quantile normalisation \citep{Bolstad03} as 
  implemented in the \Rfunction{normalize.quantiles} function of the 
  \Rpackage{preprocessCore} package. 
\item{\Robject{quantiles.robust}} Applies robust quantile normalisation 
  \citep{Bolstad03} as implemented in the 
  \Rfunction{normalize.quantiles.robust} function of the 
  \Rpackage{preprocessCore} package. 
\item{\Robject{vsn}} Applies variance stabilisation normalization \citep{Huber2002} 
  as implemented in the \Rfunction{vsn2} function of the 
  \Rpackage{vsn} package. 
\item{\Robject{max}} Each feature's reporter intensity is divided by the 
  maximum of the reporter ions intensities.
\item{\Robject{sum}} Each feature's reporter intensity is divided by the 
  sum of the reporter ions intensities.
\end{description}

<<normalise,echo=TRUE,cache=FALSE>>=
qnt.max <- normalise(qnt, "max")
qnt.sum <- normalise(qnt, "sum")
qnt.quant <- normalise(qnt, "quantiles")
qnt.qrob <- normalise(qnt, "quantiles.robust")
qnt.vsn <- normalise(qnt, "vsn")
@

The effect of these are illustrated on figure \ref{fig:norm-plot} and
figure \ref{fig:cv-plot} reproduces figure 3 of \cite{Karp2010} that described 
the application of vsn on iTRAQ reporter data.

\begin{figure}[p]
<<norm-plot, dev='tikz', fig.width=5, fig.height=7, echo=FALSE>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[2],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt, ttl = "Non-normalised data")
.plot(qnt.max, ttl = "Maximum")
.plot(qnt.sum, ttl = "Sum")
.plot(qnt.quant, ttl = "Quantile")
.plot(qnt.qrob, ttl = "Robust quantile")
.plot(qnt.vsn, ttl = "vsn")
@ 
\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}

<<cvdata,echo=FALSE,cache=FALSE>>=
sd1 <- apply(log2(exprs(qnt))+10,1,sd)
mn1 <- apply(log2(exprs(qnt))+10,1,mean)
cv1 <- sd1/mn1
sd2 <- apply(exprs(qnt.vsn)+10,1,sd)
mn2 <- apply(exprs(qnt.vsn)+10,1,mean)
cv2 <- sd2/mn2
dfr <- rbind(data.frame(rank=order(mn1),cv=cv1,norm="raw"),
             data.frame(rank=order(mn2),cv=cv2,norm="vsn"))
library("zoo")
## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
## rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
##
## Calling directly rollapply.zoo to make it zoo_1.6-4 compatible.
## The above requires zoo >= 1.7-0, which is as of 15 March 2011
## not yet available on CRAN (only on r-forge).
rmed1 <- zoo:::rollapply.zoo(cv1,7,function(x) median(x,na.rm=TRUE))
rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
dfr2 <- 
  rbind(data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed1,norm="raw"),
        data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed2,norm="vsn"))

p <- qplot(rank,cv,data=dfr,col=norm) + 
  geom_line(data=dfr2,aes(x=x,y=y,colour=norm)) +
  theme_gray(7)
@

\begin{figure}[!ht]
<<cv-plot, dev='tikz', fig.width=5, fig.height=4, echo=FALSE>>=
print(p)
@ 
\caption{CV versus signal intensity comparison for log2 and vsn
  transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

Note that it is also possible to normalise individual spectra or whole \Robject{MSnExp} experiments with 
the \Rfunction{normalise} method using the \Robject{max} method. 
This will rescale all peaks between 0 and 1. 
To visualise the relative reporter peaks, one should this first trim the spectra 
using method \Rfunction{trimMz} as illustrated in section \ref{sec:processing}, 
then normalise the \Robject{MSnExp} with \Rfunction{normalise} using 
\Robject{method="max"} as illustrated above and plot the data using \Rfunction{plot} 
(figure \ref{fig:msnexp-norm-plot}).

<<prepare-msnset-nrom-plot, cache=FALSE, echo=FALSE, keep.fig='none'>>=
p <- plot(normalise(experiment[bsasel], "max"),
          reporters = iTRAQ4, full = FALSE, plot = FALSE)
p <- p + theme_gray(7)
@

\begin{figure}[p]
<<msnexp-norm-plot, dev='tikz', fig.width=5, fig.height=6, echo=FALSE>>=
print(p)
@ 
\caption{Experiment-wide normalised MS2 spectra. The y-axes of the individual 
  spectra is now rescaled between 0 and 1 (highest peak), as opposed to 
  figure \ref{fig:msnexp-plot}. }
\label{fig:msnexp-norm-plot}
\end{figure}

\subsection{Feature aggregation}
\label{sec:feataggregation}

The above quantitation and normalisation has been performed on quantitative data 
obtained from individual spectra. However, the biological unit of interest is not 
the spectrum but the peptide or the protein. As such, it is important to be able to 
summarise features that belong to a same group, i.e. spectra from one peptide, 
peptides that originate from one protein, or directly combine all spectra that have 
been uniquely associated to one protein.

\Rpackage{MSnbase} provides one function, \Rfunction{combineFeatures}, that allows 
to aggregate features stored in an \Robject{MSnSet} using build-in or user 
defined summary function and return a new \Robject{MSnSet} instance. 
The three main arguments are described below. 
Additional details can be found in the method documentation.

<<makeGroups1,echo=FALSE,cache=FALSE>>=
gb <- fData(qnt)$ProteinAccession
@

\Rfunction{combineFeatures}'s first argument, \Robject{object}, is an instance of class \Robject{MSnSet}, 
as has been created in the section \ref{sec:quant} for instance. 
The second argument, \Robject{groupBy}, is a \Robject{factor} than has as many elements as 
there are features in the \Robject{MSnSet} \Robject{object} argument. The features corresponding 
to the \Robject{groupBy} levels will be aggregated so that the resulting \Robject{MSnSet} output 
will have \Rfunction{length(levels(groupBy))} features. Here, we will combine individual 
MS2 spectra based on the protein they originate from. 
As shown below, this will result in \Sexpr{length(table(gb))} new aggregated features.

<<makeGroups2,echo=TRUE,cache=FALSE>>=
gb <- fData(qnt)$ProteinAccession
table(gb)
length(unique(gb))
@

The third argument, \Robject{fun}, 
defined how to combine the features. Predefined functions are readily available and can be 
specified as strings 
(\Robject{fun="mean"}, \Robject{fun="median"}, \Robject{fun="sum"}, 
\Robject{fun="weighted.mean"} or \Robject{fun="medianpolish"} to compute respectively the mean, 
media, sum, weighted mean or median polish of the features to be aggregated). 
Alternatively, is is possible to supply user defined functions with 
\Robject{fun=function(x) \{ ... \}}. We will use the \Robject{median} here.

<<combineFeatures,echo=TRUE,cache=FALSE>>=
qnt2 <- combineFeatures(qnt, groupBy = gb, fun = "median")
qnt2
@

%% Other intended ways to define aggregation groups is to import 
%% identification data into R and collate it to the \Robject{MSnSet} 
%% feature data, and subsequently use peptide sequences or protein assignments 
%% to combine features.

\subsection{Label-free MS2 quantitation}
\label{sec:spectralcounting}

Note that if samples are not multiplexed, label-free MS2 quantitation is possible 
using \Rpackage{MSnbase}.
Once individual spectra have been assigned to peptides and proteins, it becomes 
straightforward to estimate protein quantities using the spectral counting method, 
as illustrated in section \ref{sec:feataggregation}, when the \Robject{groupBy} 
argument is defined.

\section{Quantitative assessment of incomplete dissociation}
\label{sec:incompdissoc}

Quantitation using isobaric reporter tags assumes complete dissociation between
the reporter group (red on figure \ref{fig:itraqchem}), 
balance group (blue) and peptide (the peptide reactive group is drawn in green). 
However, incomplete dissociation does occur and 
results in an isobaric tag (i.e reporter and balance groups) specific peaks.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=4cm]{./itraqchem.pdf}
\caption{ iTRAQ 4-plex isobaric tags reagent consist of three parts: 
  (1) a charged reporter group (MZ of 114, 115, 116 and 117) that is unique to each of the four reagents (red),
  (2) an uncharged mass balance group (28-31 Da) (blue)and 
  (3) a peptide reactive group (NHS ester) that binds to the peptide. 
  In case of incomplete dissociation, the reporter and balance groups produce a specific peaks 
  at MZ 145.
}
\label{fig:itraqchem}
  \end{center}  
\end{figure}

\Rpackage{MSnbase} provides, among others, a \Robject{ReporterIons} object for iTRAQ 4-plex that includes the 
145 peaks, called \Robject{iTRAQ5}. This can then be used to quantify the experiment as show in section \ref{sec:quant} 
to estimate incomplete dissociation for each spectrum.

<<incompdiss,echo=TRUE,cache=FALSE,keep.source=T>>=
iTRAQ5
incompdiss <- quantify(itraqdata,
                       method = "trap",
                       reporters = iTRAQ5,
                       strict = FALSE,
                       parallel = FALSE,
                       verbose = FALSE)
head(exprs(incompdiss))
@

Figure \ref{fig:incompdiss} compares these intensities for the whole experiment.

\begin{figure}[!hbt]
<<incompdiss-plot, dev='tikz', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
dfr <- melt(exprs(incompdiss))
colnames(dfr) <- c("feature", "reporters", "intensity")
dfr$reporters <- sub("iTRAQ5.", "", dfr$reporters)
repsum <- apply(exprs(incompdiss), 1, function(x) sum(x[1:4]))
dfr2 <- data.frame(iTRAQ1to4 = repsum, iTRAQ5 = exprs(incompdiss)[,5])
p1 <- ggplot(data = dfr, aes(x = reporters,y = log10(intensity))) +
  geom_boxplot() +
  theme_gray(6)
p2 <- ggplot(data = dfr2, aes(x = log10(iTRAQ1to4), y = log10(iTRAQ5))) +
  geom_point(alpha = I(1/2)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  stat_smooth(method = lm, se = FALSE) +
  xlab(label = expression(log[10]~sum~114~to~117)) +
  ylab(label = expression(log[10]~145)) +
  theme_gray(6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
print(p1, vp = vplayout(1, 1))
print(p2, vp = vplayout(1, 2))
@ 
\caption{Boxplot and scatterplot comparing intensities of the 4 reporter ions 
  (or their sum, on the right) and the incomplete dissociation specific peak. }
\label{fig:incompdiss}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Combining \Robject{MSnSet} instances}\label{sec:combine}

Combining mass spectrometry runs can be done in two different ways depending on the 
nature of these runs. If the runs represent repeated measures of identical samples, 
for instance multiple fractions, the data has to be combine along the row of the 
quantitation matrix: all the features (along the rows) represent measurements of the 
same set of samples (along the columns). In this situation, described in section 
\ref{sec:comb1}, two experiments of dimensions 
$n_1$ (rows) by $m$ (columns and $n_2$ by $m$ will produce 
a new experiment of dimensions $n_1 + n_2$ by $m$.

When however, different sets of samples have been analysed in different mass 
spectrometry runs, the data has 
to be combined along the columns of the quantitation matrix: some features will be 
shared across experiments and should thus be aligned on a same row in the new
data set, whereas unique 
features to one experiment should be set as missing in the other one. 
In this situation, described in section \ref{sec:comb2}, 
two experiments of dimensions $n_1$ by $m_1$ and $n_2$ by $m_2$ will 
produce a new experiment of dimensions 
$unique_{n_1} + unique_{n_2} + shared_{n_1, n_2}$ by $m_1 + m_2$. 
The two first terms of the first dimension will be complemented by \Robject{NA} values.

Default \Robject{MSnSet} feature names (\Robject{X1}, \Robject{X2}, \ldots) and sample names 
(\Robject{iTRAQ4.114}, \Robject{iTRAQ4.115}, \Robject{iTRAQ4.116}, \ldots) are not informative.
The features and samples of these anonymous quantitative data-sets should be updated 
before being combined, to guide how to meaningfully merge them.

\subsection{Combining identical samples}\label{sec:comb1}

To simulate this situation, let us use quantiation data from the \Robject{itraqdata} 
object that is provided with the package as experiment 1 and the data from the 
\Robject{rawdata} \Robject{MSnExp} instance created at the very beginning of this 
document. Both experiments share the \emph{same} default iTRAQ 4-plex reporter names 
as default sample names, and will thus automatically be combined along rows.

<<makeexp12, echo=TRUE, cache=FALSE>>=
exp1 <- quantify(itraqdata, reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
sampleNames(exp1)
exp2 <- quantify(rawdata, reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
sampleNames(exp2)
@

It important to note that the features of these independent experiments share 
the same default anonymous names: X1, X2, X3, \ldots, 
that however represent quantitation of distinct physical analytes. 
If the experiments were to be combined as is, it would result in an error because
data points for the same \emph{feature} name (say \Robject{X1}) and the 
same \emph{sample name} (say \Robject{iTRAQ4.114}) have different values. 
We thus first update the feature names to explicitate that they originate from 
different experiment and represent quantitation from different spectra 
using the convenience function \Rfunction{updateFeatureNames}.
Note that updating the names of one experiment would suffice here.

<<updateFnames, echo=TRUE, cache=FALSE>>=
head(featureNames(exp1))
exp1 <- updateFeatureNames(exp1)
head(featureNames(exp1))
head(featureNames(exp2))
exp2 <- updateFeatureNames(exp2)
head(featureNames(exp2))
@

The two experiments now share the same sample names and have different feature names 
and will be combined along the row. Note that all meta-data is correctly 
combined along the quantitation values.

<<comb1, echo=TRUE, cache=FALSE>>=
exp12 <- combine(exp1, exp2)
dim(exp1)
dim(exp2)
dim(exp12)
@

\subsection{Combine different samples}\label{sec:comb2}

Lets now create two \Robject{MSnSet}s from the same raw data to simulate two different 
independent experiments that share some features. 
As done previously (see section \ref{sec:feataggregation}), 
we combine the specta based on the proteins they have been identified to belong to.
Features can thus naturally be named using protein accession numbers.
Alternatively, if peptide sequences would have been used as grouping factor in 
\Rfunction{combineFeatures}, then these would be good feature name candidates.

<<make2exps2, echo=TRUE, cache=FALSE>>=
set.seed(1)
i <- sample(length(itraqdata), 35)
j <- sample(length(itraqdata), 35)
exp1 <- quantify(itraqdata[i], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
exp2 <- quantify(itraqdata[j], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
table(featureNames(exp1) %in% featureNames(exp2))
exp1 <- combineFeatures(exp1,
                        groupBy = fData(exp1)$ProteinAccession)
exp2 <- combineFeatures(exp2,
                        groupBy = fData(exp2)$ProteinAccession)
head(featureNames(exp1))
head(featureNames(exp1))
@

The feature names are updated automatically by \Rfunction{combineFeatures}, using the \Robject{groupBy} argument. 
Proper feature names, reflecting the nature of the features (spectra, peptides or proteins) is critical when multiple experiments are to be combined, as this is done using common features as defined by their names (see below).

\bigskip

Sample names should also be updated to replace anonymous reporter names with relevant identifiers; 
the individual reporter data is stored in the \Robject{phenoData} and is not lost.
A convenience function \Rfunction{updateSampleNames} is provided to append the \Robject{MSnSet}'s 
variable name to the already defined names, although in general, biologically relevant 
identifiers are preferred.

<<renameSamples, echo=TRUE, cache=FALSE>>=
sampleNames(exp1)
exp1 <- updateSampleNames(exp1)
sampleNames(exp1)
sampleNames(exp1) <- c("Ctrl1", "Cond1", "Ctrl2", "Cond2")
sampleNames(exp2) <- c("Ctrl3", "Cond3", "Ctrl4", "Cond4")
@

At this stage, it is not yet possible to combine the two experiments, 
because their feature data is not compatible yet; they share the same feature 
variable labels, i.e. the feature data column names 
(\Sexpr{paste(head(fvarLabels(exp1), n = 3), collapse=", ")}, \ldots), 
but the part of the content is different because the original data was (in particular 
all the spectrum centric data: identical peptides in different runs will have different 
retention times, precursor intensities, \ldots). 
Feature data with identical labels (columns in the data frame) 
and names (row in the data frame) are expected to have the same data and produce an
error if not conform.

<<fdatanames, echo=TRUE, cache=FALSE>>=
stopifnot(all(fvarLabels(exp1) == fvarLabels(exp2)))
fData(exp1)["BSA", 1:4]
fData(exp2)["BSA", 1:4]
@

Instead of removing these identical feature data columns, 
one can use a second convenience function, \Rfunction{updateFvarLabels}, 
to update feature labels based on the experiements variable name 
and maintain all the metadata.

<<renameFvars, echo=TRUE, cache=FALSE>>=
exp1 <- updateFvarLabels(exp1)
exp2 <- updateFvarLabels(exp2)
head(fvarLabels(exp1))
head(fvarLabels(exp2))
@

It is now possible to combine \Robject{exp1} and \Robject{exp2}, 
including all the meta-data, with the \Rfunction{combine} method. 
The new experiment will contain the union of 
the feature names of the individual experiments with missing values 
inserted appropriately.

<<combine, echo=TRUE, cache=FALSE>>=
exp12 <- combine(exp1, exp2)
dim(exp12)
pData(exp12)
exprs(exp12)[25:28, ]
exp12
@

\bigskip

In summary, when experiments with different samples need to be combined 
(along the columns), one needs to 
(1) clarify the sample names using \Rfunction{updateSampleNames} or
better manually, for biological relevance and (2) update the feature 
data variable labels with \Rfunction{updateFvarLabels}. 
The individual experiments (there can be more than 2) can then easily be combined 
with the \Rfunction{combine} method while retaining the meta-data.

If runs for the same sample (different fractions for example) need to be combines, 
one needs to (1) differentiate the feature provenance with 
\Rfunction{updateFeatureNames} prior to use \Rfunction{combine}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \section{Conclusion}\ref{sec:ccl}

\section{Session information}\label{sec:sessionInfo} 
<<label='sessioninfo', results=tex, echo=FALSE, cache=FALSE>>=
toLatex(sessionInfo())
@

\bibliographystyle{plainnat}
\bibliography{MSnbase}

\end{document}