### 2016-12-18 docs/articles/benchmarking.Rmd
udpate docs

---
title: "MSnbase2 benchmarking"
output:
  BiocStyle::html_document2:
     toc: false
vignette: >
  % \VignetteIndexEntry{MSnbase benchmarking}
  % \VignetteEngine{knitr::rmarkdown}
  % \VignetteKeyword{proteomics, mass spectrometry}
  % \VignettePackage{MSnbase}
---

```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
```

```{r env, echo=FALSE, message=FALSE}
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("BiocParallel"))
```

**Package**: `r Biocpkg("MSnbase")`<br />
**Authors**: Laurent Gatto and Johannes Rainer<br />
**Modified:** `r file.info("benchmarkding.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

In this vignette, we will document various timings and benchmarkings
of the recent `r Biocpkg("MSnbase")` development (aka `MSnbase2`),
that focuses on *on-disk* data access (as opposed to
*in-memory*). More details about the new implementation will be
documented elsewhere.

As a benchmarking dataset, we are going to use a subset of an TMT
6-plex experiment acquired on an LTQ Orbitrap Velos, that is
distributed with the `r Biocexptpkg("msdata")` package

```{r msdata}
library("msdata")
f <- msdata::proteomics(full.names = TRUE, pattern = "TMT_Erwinia")
basename(f)
```

We need to load the `r Biocpkg("MSnbase")` package and set the
session-wide verbosity flag to `FALSE`.

```{r verb}
library("MSnbase")
setMSnbaseVerbose(FALSE)
```

# Reading data

We first read the data using the original `readMSData` function that
generates an in-memory representation of the MS2-level raw data and
measure the time needed for this operation.

```{r read1}
system.time(inmem <- readMSData(f, msLevel = 2,
                                centroided = TRUE))
```

Next, we use the `readMSData2` function to generate an on-disk
representation of the same data.

```{r read2}
system.time(ondisk <- readMSData2(f, msLevel = 2,
                                  centroided = TRUE))
```

Creating the on-disk experiment is considerable faster and scales to
much bigger, multi-file data, both in terms of object creation time,
but also in terms of object size (see next section). We must of course
make sure that these two datasets are equivalent:

```{r equal12}
all.equal(inmem, ondisk)
```

# Data size

To compare the size occupied in memory of these two objects, we are
going to use the `object_size` function from the `r CRANpkg("pryr")`
package, which accounts for the data (the spectra) in the `assayData`
environment (as opposed to the `object.size` function from the `utils`
package).

```{r}
library("pryr")
object_size(inmem)
object_size(ondisk)
```

The difference is explained by the fact that for `ondisk`, the spectra
are not created and stored in memory; they are access on disk when
needed, such as for example for plotting:

```{r plot0, eval=FALSE}
plot(inmem[[200]], full = TRUE)
plot(ondisk[[200]], full = TRUE)
```

```{r plot1, echo=FALSE, fig.wide=TRUE, fig.cap = "Plotting in-memory and on-disk spectra"}
suppressMessages(requireNamespace("gridExtra"))
gridExtra::grid.arrange(plot(inmem[[200]],  full = TRUE),
                        plot(ondisk[[200]], full = TRUE),
                        ncol = 2)

```

# Accessing spectra

The drawback of the on-disk representation is when the spectrum data
has to actually be accessed. To compare access time, we are going to
use the `r CRANpkg("microbenchmark")` and repeat access 10 times to
compare access to all `r length(inmem)` and a single spectrum
in-memory (i.e. pre-loaded and constructed) and on-disk
(i.e. on-the-fly access).

```{r mb, cache=TRUE}
library("microbenchmark")
mb <- microbenchmark(spectra(inmem),
                     inmem[[200]],
                     spectra(ondisk),
                     ondisk[[200]],
                     times = 10)
mb
```

While it takes order or magnitudes more time to access the data on-the-fly
rather than a pre-generated spectrum, accessing all spectra is only marginally
slower than accessing all spectra, as most of the time is spent preparing the
file for access, which is done only once.

On-disk access performance will depend on the read throughput of the
disk. A comparison of the data import of the above file from an
internal solid state drive and from an USB3 connected hard disk showed
only small differences for the `readMSData2` call (1.07 *vs* 1.36
seconds), while no difference were observed for accessing individual
or all spectra. Thus, for this particular setup, performance was about
the same for SSD and HDD. This might however not apply to setting in
which data import is performed in parallel from multiple files.

<!-- ```{r ssdVsSata, eval = FALSE, echo = FALSE} -->
<!-- ## Reading the file from the internal SSD harddisk. -->
<!-- file_name <- "TMT_Erwinia_1uLSike_Top10HCD_isol2_45stepped_60min_01.mzML.gz" -->
<!-- f_1 <- paste0("/Users/jo/tmp/", file_name) -->
<!-- system.time(ondisk_1 <- readMSData2(f_1, -->
<!--                                     centroided = TRUE)) -->
<!-- ## Reading the same file from an external USB3 connected HD -->
<!-- f_2 <- paste0("/Volumes/jodata/tmp/", file_name) -->
<!-- system.time(ondisk_2 <- readMSData2(f_2, -->
<!--                                     centroided = TRUE)) -->

<!-- mb <- microbenchmark(spectra(ondisk_1), -->
<!--                      ondisk_1[[200]], -->
<!--                      spectra(ondisk_2), -->
<!--                      ondisk_2[[200]], -->
<!--                      times = 10) -->
<!-- ``` -->

Data access does not prohibit interactive usage, such as
plotting, for example, as it is about 1/2 seconds, which is an
operation that is relatively rare, compared to subsetting and
filtering, which are faster for on-disk data:

```{r subset}
i <- sample(length(inmem), 100)
system.time(inmem[i])
system.time(ondisk[i])
```

Operations on the spectra data, such as peak picking, smoothing,
cleaning, ... are cleverly cached and only applied when the data is
accessed, to minimise file access overhead. Finally, specific
operations such as for example quantitation (see next section) are
optimised for speed.

# MS2 quantitation

Below, we perform TMT 6-plex reporter ions quantitation on the first
100 spectra and verify that the results are identical (ignoring
feature names).

```{r qnt, cache=TRUE}
system.time(eim <- quantify(inmem[1:100], reporters = TMT6,
                            method = "max"))
system.time(eod <- quantify(ondisk[1:100], reporters = TMT6,
                            method = "max"))
all.equal(eim, eod, check.attributes = FALSE)
```

# Conclusions

This document focuses on speed and size improvements of the new
on-disk `MSnExp` representation. The extend of these improvements will
substantially increase for larger data.

For general functionality about the on-disk `MSnExp` data class and
`r Biocpkg("MSnbase")` in general, see other vignettes available with

```{r vigs, eval=FALSE}
vignette(package = "MSnbase")
```