### 2016-12-23 README.Rmd
Move NEWS to NEWS.md
The NEWS file is not in markdown format. Major version are first
level sections and should only comprise x.y version numbers. z-level
changes are documented as second level section starting with
'Changes in version x.y.z'.

This change is motivated by the use of 'pckdown' to generate package
webpages, which does not support anything other than NEWS.md.

[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) [![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)

```{r env, echo=FALSE}
suppressPackageStartupMessages(library("BiocStyle"))
knitr::opts_chunk$set(comment=NA)
```

`r Biocpkg("MSnbase")` is an R/Bioconductor package that provides
infrastructure for plotting, data manipulation and processing of
MS-based proteomics data. The project was started by
[Laurent Gatto](http://lgatto.github.io/) in October 2010 (Mon Oct 4
23:35:23 2010, according to the git log) and has, since then,
benefited from
[various contributions](http://lgatto.github.io/msnbase-contribs/).

The official page is the Bioconductor landing page
([release](http://www.bioconductor.org/packages/release/bioc/html/MSnbase.html)
or
[devel](http://www.bioconductor.org/packages/devel/bioc/html/MSnbase.html)
versions). The [github page](https://github.com/lgatto/MSnbase) page
is for sharing, testing, issue tracking and forking/pulling
purposes. The
[Bioconductor svn server](https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/),
is the official repository for the working source code.

## Installation

To install `r Biocpkg("MSnbase")`

```{r install, eval = FALSE}
library("BiocInstaller")
biocLite("MSnbase")
```

If you need the github version (not recommended unless you know what
you are doing)

```{r installgh, eval = FALSE}
biocLite("lgatto/MSnbase")
```

## Questions

General questions should be asked on the
[Bioconductor support forum](https://support.bioconductor.org/), using
`MSnbase` to tag the question. Feel also free to open a GitHub
[issue](https://github.com/lgatto/MSnbase/issues).

## Citation

To cite the `MSnbase` package in publications, please use:

> Gatto L, Lilley KS. *`MSnbase` - an R/Bioconductor package for
> isobaric tagged mass spectrometry data visualization, processing and
> quantitation*. Bioinformatics. 2012 Jan
> 15;28(2):288-9. doi:[10.1093/bioinformatics/btr645](http://bioinformatics.oxfordjournals.org/content/28/2/288.long). Epub
> 2011 Nov 22.  PubMed
> [PMID:22113085](http://www.ncbi.nlm.nih.gov/pubmed/22113085).

## Contributing

Contributions to the package are more than welcome. If you want to
contribute to this package, you should follow the same conventions as
the rest of the functions whenever it makes sense to do so. Feel free
to get in touch (preferable opening a
[github issue](https://github.com/lgatto/MSnbase/issues/)) to discuss
any suggestions. The
[`MSnbase` development vignette](http://lgatto.github.io/MSnbase/articles/MSnbase-development.html)
gives some background on the class infrastructure.

Please note that this project is released with a
[Contributor Code of Conduct](https://github.com/lgatto/MSnbase/blob/master/CONDUCT.md). By
participating in this project you agree to abide by its terms.