### 2016-12-23 docs/articles/Figures/MSnbase-io-out.tex
Merge branch 'master' into devel
* master: (24 commits)
  fix typos from d342c70 in README.Rmd
  update installation ignore
  fix typo [ci skip]
  udpate docs with verbose is FALSE
  set default MSnbase verbose to FALSE
  update docs
  update docs
  update news with new news format
  Move NEWS to NEWS.md
  Migrate io and dev vignettes to BiocStyle's html_document2 style
  removes notes
  mv devel vignette to BiocStyle html2
  use BiocStyle's html_docuement2 for benchmarking vignette
  Amend show method for MSnExp
  Small change to show method for MSnExp
  udpate docs
  update docs with Seb's s

\documentclass{minimal}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,shadows,fit,automata}

%%% <
\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}%
%%% >

\begin{comment}
  :Title:  MSnbase Output
  :Author: Laurent Gatto
\end{comment}

%% pgf setup
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
% Define block styles
\tikzstyle{input} = [rectangle, draw, fill=blue!20,
text width=6em, text centered, rounded corners, minimum height=4em]
\tikzstyle{fun} = [rectangle, draw, fill=white, drop shadow,
text width=7em, text centered, rounded corners, minimum height=2em]
\tikzstyle{obj} = [rectangle, draw, fill=red!20,
text width=5em, text centered, rounded corners, minimum height=5em]

\begin{document}

\begin{tikzpicture}[node distance = 2cm, auto]
  % Place nodes
  \node [input] (raw) {Raw data in an open \texttt{XML} format};
  \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
  \node [fun, right of=mgf, node distance=4cm, yshift=1cm] (writeMgfData) {\texttt{writeMgfData}};
  \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
  % \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
  \node [fun, right of=spreadsheet, node distance=4cm] (writeexprs) {\texttt{write.exprs}};
  % \node [fun, right of=mztab, node distance=4cm] (writeMzTabData) {\texttt{writeMzTabData}};
  \node [obj, right of=writeMgfData, yshift=1.25cm, node distance=4cm] (MSnExp) {\texttt{MSnExp}};
  \node [obj, right of=writeexprs, node distance=4cm] (MSnSet) {\texttt{MSnSet}};
  \node [fun, below of=MSnExp] (quantify) {\texttt{quantify}};
  % Background
  \begin{pgfonlayer}{background}
    \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
  \end{pgfonlayer}
  % Draw edges
  \draw  (MSnExp) -- (writeMgfData);
  \draw [->] (writeMgfData) -- (mgf);
  \draw (MSnSet) -- (writeexprs);
  \draw [->] (writeexprs) -- (spreadsheet);
  \draw  (MSnExp) -- (quantify);
  \draw [->] (quantify) -- (MSnSet);
  % \draw (MSnSet) -- (writeMzTabData);
  % \draw [->] (writeMzTabData) -- (mztab);
\end{tikzpicture}

\end{document}