### 2016-12-23 docs/articles/MSnbase-development.Rnw

%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{MSnbase development}
%\VignetteKeywords{Mass Spectrometry, Proteomics, Infrastructure }
%\VignettePackage{MSnbase-development}

\documentclass[12pt, oneside]{article}

<<style, eval=TRUE, echo=FALSE, results="asis">>=
BiocStyle::latex2()
@

\author{
  Laurent Gatto and Johannes Rainer
}

\bioctitle[\Biocpkg{MSnbase} devel]{\Biocpkg{MSnbase} development }

\begin{document}

\maketitle

\begin{abstract}
  This vignette describes the classes implemented in \Biocpkg{MSnbase}
  package.  It is intended as a starting point for developers or users
  who would like to learn more or further develop/extend mass
  spectrometry and proteomics data structures.
\end{abstract}

<<'setup', include = FALSE, cache = FALSE>>=
library("knitr")
opts_chunk$set(fig.align = 'center',
               fig.show = 'hold',
               par = TRUE,
               prompt = FALSE,
               comment = NA)
options(replace.assign = TRUE,
        width = 65)
@

<<environment, cache=FALSE, echo=FALSE>>=
suppressPackageStartupMessages(library(MSnbase))
@

\newpage

\tableofcontents

\newpage

\newpage

\input{Foreword.tex}

\input{Bugs.tex}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

This document is not a replacement for the individual manual pages, that
document the slots of the \Biocpkg{MSnbase} classes. It is a centralised high-level
description of the package design.

\Biocpkg{MSnbase} aims at being compatible with the \Biocpkg{Biobase}
infrastructure \cite{Gentleman2004}.  Many meta data structures that
are used in \Rclass{eSet} and associated classes are also used
here. As such, knowledge of the \textit{Biobase development and the
  new eSet} vignette would be beneficial; the vignette can directly be
accessed with
\texttt{vignette("BiobaseDevelopment", package="Biobase")}.

The initial goal is to use the \Biocpkg{MSnbase} infrastructure for
MS$^2$ labelled (iTRAQ \cite{Ross2004} and TMT \cite{Thompson2003})
and label-free (spectral counting, index and abundance ) quantitation
- see the documentation for the \Rfunction{quantify} function for
details.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\Biocpkg{MSnbase} classes}

All classes have a \Rclass{.\_\_classVersion\_\_} slot, of class
\Rclass{Versioned} from the \Biocpkg{Biobase} package. This slot
documents the class version for any instance to be used for debugging
and object update purposes. Any change in a class implementation
should trigger a version change.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{pSet}: a virtual class for raw mass spectrometry data and meta data}
\label{sec:pset}

This virtual class is the main container for mass spectrometry data,
i.e spectra, and meta data. It is based on the \Rclass{eSet}
implementation for genomic data. The main difference with
\Rclass{eSet} is that the \Robject{assayData} slot is an environment
containing any number of \Rclass{Spectrum} instances (see section
\ref{sec:spectrum}).

One new slot is introduced, namely \Robject{processingData}, that
contains one \Rclass{MSnProcess} instance (see section
\ref{sec:msnprocess}).  and the \Robject{experimentData} slot is now
expected to contain \Rclass{MIAPE} data (see section \ref{sec:miape}).
The \Robject{annotation} slot has not been implemented, as no prior
feature annotation is known in shotgun proteomics.

<<pSet>>=
getClass("pSet")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{MSnExp}: a class for MS experiments}
\label{sec:msnexp}

\Rclass{MSnExp} extends \Rclass{pSet} to store MS experiments.
It does not add any new slots
to \Rclass{pSet}. Accessors and setters are all inherited from
\Rclass{pSet} and new ones should be implemented for \Rclass{pSet}.
Methods that manipulate actual data in experiments are implemented
for \Rclass{MSnExp} objects.

<<MSnExp>>=
getClass("MSnExp")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{OnDiskMSnExp}: a on-disk implementation of the \Rclass{MSnExp} class}
\label{sec:ondiskmsnexp}

The \Rclass{OnDiskMSnExp} class extends \Rclass{MSnExp} and inherits
all of its functionality but is aimed to use as little memory as
possible based on a balance between memory demand and
performance. Most of the spectrum-specific data, like retention time,
polarity, total ion current are stored within the object's
\Robject{featureData} slot. The actual M/Z and intensity values from
the individual spectra are, in contrast to \Rclass{MSnExp} objects,
not kept in memory (in the \Robject{assayData} slot), but are fetched
from the original files on-demand. Because mzML files are indexed,
using the \Biocpkg{mzR} package to read the relevant spectrum data is
fast and only moderately slower than for in-memory
\Rclass{MSnExp}\footnote{The \textit{benchmarking} vignette compares
  data size and operation speed of the two implementations.}.

To keep track of data manipulation steps that are applied to spectrum
data (such as performed by methods \Rfunction{removePeaks} or
\Rfunction{clean}) a \textit{lazy execution} framework was
implemented. Methods that manipulate or subset a spectrum's M/Z or
intensity values can not be applied directly to a
\Rclass{OnDiskMSnExp} object, since the relevant data is not kept in
memory. Thus, any call to a processing method that changes or subset
M/Z or intensity values are added as \Rclass{ProcessingStep} items to
the object's \Robject{spectraProcessingQueue}. When the spectrum data
is then queried from an \Rclass{OnDiskMSnExp}, the spectra are read in
from the file and all these processing steps are applied on-the-fly to
the spectrum data before being returned to the user.

% This has obviously a negative impact on the performance (e.g. for
% repeated function calls), would however also allow to remove certain
% processing steps at a later stage.

The operations involving extracting or manipulating spectrum data are
applied on a per-file basis, which enables parallel processing. Thus,
all corresponding method implementations for \Rclass{OnDiskMSnExp}
objects have an argument \Robject{BPPARAM} and users can set a
\Rcode{PARALLEL\_THRESH} option flag \footnote{see
  \Rfunction{?MSnbaseOptions} for details.} that enables to define how
and when parallel processing should be performed (using the
\Biocpkg{BiocParallel} package).

Note that all data manipulations that are not applied to M/Z or
intensity values of a spectrum (e.g. sub-setting by retention time
etc) are very fast as they operate directly to the object's
\Rclass{featureData} slot.

<<OnDiskMSnExp>>=
getClass("OnDiskMSnExp")
@

The distinction between \Rclass{MSnExp} and \Rclass{OnDiskMSnExp} is
often not explicitly stated as it should not matter, from a user's
perspective, which data structure they are working with, as both
behave in equivalent ways. Often, they are referred to as
\textit{in-memory} and \textit{on-disk} \Rclass{MSnExp}
implementations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{MSnSet}: a class for quantitative proteomics data}
\label{sec:msnset}

This class stores quantitation data and meta data after running
\Rfunction{quantify} on an \Rclass{MSnExp} object or by creating an
\Rclass{MSnSet} instance from an external file, as described in the
\texttt{MSnbase-io} vignette and in \Rfunction{?readMSnSet},
\Rfunction{readMzTabData}, etc. The quantitative data is in form of a
$m \times n$ matrix, where $m$ is the number of features/spectra
originally in the \Rclass{MSnExp} used as parameter in
\Rfunction{quantify} and $m$ is the number of reporter ions (see
section \ref{sec:reporterions}). If read from an external file, $n$
corresponds to the number of features (protein groups, proteins,
peptides, spectra) in the file and $m$ is the number of columns with
quantitative data (samples) in the file.

This prompted to keep a similar implementation as the
\Rclass{ExpressionSet} class, while adding the proteomics-specific
annotation slot introduced in the \Rclass{pSet} class, namely
\Robject{processingData} for objects of class \Rclass{MSnProcess} (see
section \ref{sec:msnprocess}).

<<MSnSet>>=
getClass("MSnSet")
@

The \Rclass{MSnSet} class extends the virtual \Rclass{eSet} class to
provide compatibility for \Rclass{ExpressionSet}-like behaviour.  The
experiment meta-data in \Robject{experimentData} is also of class
\Rclass{MIAPE} (see section \ref{sec:miape}).  The
\Robject{annotation} slot, inherited from \Rclass{eSet} is not
used. As a result, it is easy to convert \Rclass{ExpressionSet} data
from/to \Rclass{MSnSet} objects with the coersion method
\Rfunction{as}.

<<as>>=
data(msnset)
class(msnset)
class(as(msnset, "ExpressionSet"))

data(sample.ExpressionSet)
class(sample.ExpressionSet)
class(as(sample.ExpressionSet, "MSnSet"))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{MSnProcess}: a class for logging processing meta data}
\label{sec:msnprocess}

This class aims at recording specific manipulations applied to
\Rclass{MSnExp} or \Rclass{MSnSet} instances. The \Robject{processing}
slot is a \Robject{character} vector that describes major
processing. Most other slots are of class \Robject{logical} that
indicate whether the data has been centroided, smoothed, \ldots
although many of the functionality is not implemented yet.  Any new
processing that is implemented should be documented and logged here.

It also documents the raw data file from which the data originates
(\Robject{files} slot) and the \Biocpkg{MSnbase} version that was in
use when the \Rclass{MSnProcess} instance, and hence the
\Rclass{MSnExp}/\Rclass{MSnSet} objects, were originally created.

<<MSnProcess>>=
getClass("MSnProcess")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{MIAPE}: Minimum Information About a Proteomics Experiment}
\label{sec:miape}

The Minimum Information About a Proteomics
Experiment \cite{Taylor2007, Taylor2008} \Rclass{MIAPE} class
describes the experiment, including contact details,
information about the mass spectrometer and control
and analysis software.

<<MIAPE>>=
getClass("MIAPE")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{Spectrum} \textit{et al.}: classes for MS spectra}
\label{sec:spectrum}

\Rclass{Spectrum} is a virtual class that defines
common attributes to all types of spectra. MS1 and MS2 specific attributes
are defined in the \Rclass{Spectrum1} and \Rclass{Spectrum2} classes,
that directly extend \Rclass{Spectrum}.

%% The choices of attributes has been dictated by the \Rfunction{xcms:::rampRawData}
%% and \Rfunction{xcms:::rampRawDataMSn} functions and what data from the
%% \texttt{mzXML} file they gave access to.
%% It is expected that some hopefully minor changes might
%% come up here when migrating to other data import packages, that allow random access
%% to \texttt{mzXML} data and support \texttt{mzML} \cite{Martens2010}.

<<Spectrum>>=
getClass("Spectrum")
@

<<Spectrum1>>=
getClass("Spectrum1")
@

<<Spectrum2>>=
getClass("Spectrum2")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{ReporterIons}: a class for isobaric tags}
\label{sec:reporterions}

The iTRAQ and TMT (or any other peak of interest)
are implemented \Rclass{ReporterIons} instances,
that essentially defines an expected MZ position for the peak and a width around
this value as well a names for the reporters.

<<ReporterIons>>=
getClass("ReporterIons")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\Rclass{NAnnotatedDataFrame}: multiplexed \Rclass{AnnotatedDataFrame}s}
\label{sec:nannotateddataframe}

The simple expansion of the \Rclass{AnnotatedDataFrame}
classes adds the \Robject{multiplex} and \Robject{multiLabel} slots to
document the number and names of multiplexed samples.

<<NAnnotatedDF>>=
getClass("NAnnotatedDataFrame")
@

\subsection{Other classes}\label{sec:other}

\subsubsection*{Lists of \Rclass{MSnSet} instances}\label{sec:msnl}

When several \Rclass{MSnSet} instances are related to each other and
should be stored together as different objects, they can be grouped as
a list into and \Rclass{MSnSetList} object. In addition to the actual
\Rclass{list} slot, this class also has basic logging functionality
and enables iteration over the \Rclass{MSnSet} instances using a
dedicated \Rfunction{lapply} methods.

<<msl>>=
getClass("MSnSetList")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Miscellaneous}

\paragraph{Unit tests} \Biocpkg{MSnbase} implements unit tests with
the \CRANpkg{testthat} package.

\paragraph{Processing methods} Methods that process raw data, i.e. spectra
should be implemented for \Rclass{Spectrum} objects first and then
\texttt{eapply}'ed (or similar) to the \Robject{assayData} slot of an
\Rclass{MSnExp} instance in the specific method.

%% \paragraph{Why this MIAME slot?} \label{misc:whymiame} While it
%% would have been possible to transfer all data stored in
%% \Biocpkg{Biobase}'s \Rclass{MIAME} to a new \Rclass{MIAPE} class
%% and use the latter for \Robject{experimentData} slots in the
%% \Rclass{pSet} class, it would not have been possible to directly
%% transfer this to \Rclass{MSnSet} instances, as \Rclass{MSnSet}
%% classes directly inherit from the \Rclass{ExpressionSet}, whose
%% \Robject{experimentData} slot must be of class \Rclass{MIAME}.

%% \input{NoteAboutSpeedAndMemory.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Session information}
\label{sec:sessionInfo}

<<sessioninfo, results='asis', echo=FALSE>>=
toLatex(sessionInfo())
@

\bibliography{MSnbase}

\end{document}