### 2017-01-04 README.Rmd
merged with master
From: Laurent <lg390@cam.ac.uk>

git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@125626 bc3139a8-67e5-0310-9ffc-ced21a209358

[![Build Status](https://travis-ci.org/lgatto/MSnbase.svg?branch=master)](https://travis-ci.org/lgatto/MSnbase) 
[![codecov.io](https://codecov.io/github/lgatto/MSnbase/coverage.svg?branch=master)](https://codecov.io/github/lgatto/MSnbase?branch=master)
[![Bioconductor devel build Status](http://bioconductor.org/shields/build/devel/bioc/MSnbase.svg)](http://bioconductor.org/packages/devel/bioc/html/MSnbase.html) 
[![Test coverage](https://bioconductor.org/shields/coverage/devel/MSnbase.svg)](https://codecov.io/github/Bioconductor-mirror/MSnbase/branch/master)

```{r env, echo=FALSE}
suppressPackageStartupMessages(library("BiocStyle"))
knitr::opts_chunk$set(comment=NA)
```

[**MSnbase**](http://lgatto.github.io/MSnbase/) is an R/Bioconductor
package that provides infrastructure for plotting, data manipulation
and processing mass spectrometry and proteomics data. The project was
started by [Laurent Gatto](http://lgatto.github.io/) in October 2010
(Mon Oct 4 23:35:23 2010, according to the git log) and has, since
then, benefited from [various
contributions](http://lgatto.github.io/msnbase-contribs/).

The official page is the Bioconductor landing page
([release](http://www.bioconductor.org/packages/release/bioc/html/MSnbase.html)
or
[devel](http://www.bioconductor.org/packages/devel/bioc/html/MSnbase.html)
versions). The [github page](https://github.com/lgatto/MSnbase) page
is for sharing, testing, issue tracking and forking/pulling
purposes. The
[Bioconductor svn server](https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/),
is the official repository for the working source code.

## Installation

To install `r Biocpkg("MSnbase")`

```{r install, eval = FALSE}
library("BiocInstaller")
biocLite("MSnbase")
```

If you need the github version (not recommended unless you know what
you are doing)

```{r installgh, eval = FALSE}
biocLite("lgatto/MSnbase")
```

## Questions

General questions should be asked on the
[Bioconductor support forum](https://support.bioconductor.org/), using
`MSnbase` to tag the question. Feel also free to open a GitHub
[issue](https://github.com/lgatto/MSnbase/issues).

## Citation

To cite the `MSnbase` package in publications, please use:

> Gatto L, Lilley KS. *`MSnbase` - an R/Bioconductor package for
> isobaric tagged mass spectrometry data visualization, processing and
> quantitation*. Bioinformatics. 2012 Jan
> 15;28(2):288-9. doi:10.1093/bioinformatics/btr645. PubMed
> [PMID:22113085](http://www.ncbi.nlm.nih.gov/pubmed/22113085).

## Contributing

Contributions to the package are more than welcome. If you want to
contribute to this package, you should follow the same conventions as
the rest of the functions whenever it makes sense to do so. Feel free
to get in touch (preferable opening a
[github issue](https://github.com/lgatto/MSnbase/issues/)) to discuss
any suggestions. The
[`MSnbase` development vignette](http://lgatto.github.io/MSnbase/articles/MSnbase-development.html)
gives some background on the class infrastructure.

Please note that this project is released with a
[Contributor Code of Conduct](https://github.com/lgatto/MSnbase/blob/master/CONDUCT.md). By
participating in this project you agree to abide by its terms.