### 2017-04-18 docs/articles/Figures/MSnbase-io-in.tex
add docs dir

\documentclass{minimal}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,shadows,fit,automata}

%%% <
\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}%
%%% >

\begin{comment}
  :Title:  MSnbase Input
  :Author: Laurent Gatto
\end{comment}

%% pgf setup
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
% Define block styles
\tikzstyle{input} = [rectangle, draw, fill=blue!20,
text width=6em, text centered, rounded corners, minimum height=4em]
\tikzstyle{fun} = [rectangle, draw, fill=white, drop shadow,
text width=7em, text centered, rounded corners, minimum height=2em]
\tikzstyle{obj} = [rectangle, draw, fill=red!20,
text width=5em, text centered, rounded corners, minimum height=5em]

\begin{document}

\begin{tikzpicture}[node distance = 2cm, auto]
  % Place nodes
  \node [input] (raw) {Raw data in an open \texttt{XML} format};
  \node [fun, left of=raw, node distance=4cm] (readMSData) {\texttt{readMSData}};
  \node [input, below of=raw] (mgf) {Peak list in \texttt{mgf} format};
  \node [fun, left of=mgf, node distance=4cm] (readMgfData) {\texttt{readMgfData}};
  \node [input, below of=mgf] (spreadsheet) {Quantitation data as a spreadsheet};
  \node [input, below of=spreadsheet] (mztab) {\texttt{mzTab} format};
  \node [fun, left of=spreadsheet, node distance=4cm] (readMSnSet) {\texttt{readMSnSet}};
  \node [fun, left of=mztab, node distance=4cm] (readMzTabData) {\texttt{readMzTabData}};
  \node [obj, left of=readMSData, yshift=-2.5mm, node distance=4cm] (MSnExp) {\texttt{MSnExp}};
  \node [obj, left of=readMSnSet, node distance=4cm] (MSnSet) {\texttt{MSnSet}};
  \node [fun, below of=MSnExp] (quantify) {\texttt{quantify}};
  % Background
  \begin{pgfonlayer}{background}
    \node [fill=yellow!20,rounded corners, draw=black!50, dashed, fit=(MSnExp) (quantify) (MSnSet)] {};
  \end{pgfonlayer}
  % Draw edges
  \draw  (raw) -- (readMSData);
  \draw  (mgf) -- (readMgfData);
  \draw  (spreadsheet) -- (readMSnSet);
  \draw [->] (readMSData) -- (MSnExp);
  \draw [->] (readMgfData) -- (MSnExp);
  \draw [->] (readMSnSet) -- (MSnSet);
  \draw  (MSnExp) -- (quantify);
  \draw [->] (quantify) -- (MSnSet);
  \draw (mztab) -- (readMzTabData);
  \draw [->] (readMzTabData) -- (MSnSet);
\end{tikzpicture}

\end{document}